BBCH data analysis for MPI trials
========================================================

## Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_phenotypes")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

## Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(ggplot2)
library(reshape)
library(pander)
library(plyr)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("BBCH.RData")
```


## Source R functions
```{r source R functions, include=FALSE}
source("../libpurzel/func_get_bbch_data.R")
source("../libpurzel/func_modify_bbch_data.R")
source("../libpurzel/func_bbch_plots.R")
source("../libpurzel/colors.R")
source("../libpurzel/names.R")
source("../libpurzel/func_anova_v2.R")
source("../libpurzel/func_aggregate_values.R")
source("../libpurzel/func_wilcoxon_test.R")
```


## Execute Query from `D:\work\TROST\Database\2014-07-29-bbch_query.sql` to get information about BBCH and meta data
### BBCH data of four TROST experiments (MPI Test1.2, MPI Test2, MPI Feld 2011, MPI Feld 2012)
```{r get data from trost database}
# use old version of function
# bbch_trost_mpi_48656 <- func_get_bbch_data_old('48656')

bbch_trost_mpi <- func_get_bbch_data( experiment_id = c('48656', '51790', '44443', '56726') )

pander(head(bbch_trost_mpi))
dim(bbch_trost_mpi)
# 3842 11
```

## Execute Query from `D:\work\TROST\Database\2015-05-07-bbch_query_valdis.sql` to get information about BBCH and meta data
### Get BBCH data for VALDIS experiments at MPI FGH 2014
```{r get BBCH data for VALDIS experiments at MPI FGH 2014}
bbch_mpi_fgh_2014 <- func_get_bbch_data2( experiment_id = '67199')

pander(head(bbch_mpi_fgh_2014))
dim(bbch_mpi_fgh_2014)
# 2535 10
```


### Modification of BBCH dataset
```{r modification of BBCH dataset}
# TROST data
bbch_trost_mpi_mod <- func_modify_bbch_data(bbch_trost_mpi)
bbch_trost_mpi_mod$cultivar <- as.factor(bbch_trost_mpi_mod$cultivar)
pander(table(bbch_trost_mpi_mod$experiment_name, bbch_trost_mpi_mod$timepoint))

# VALDIS data
bbch_mpi_fgh_2014_mod <- func_modify_bbch_data(bbch_mpi_fgh_2014)

# timepoints
table(bbch_mpi_fgh_2014_mod$date)
pander(head(table(bbch_mpi_fgh_2014_mod$line_id, bbch_mpi_fgh_2014_mod$date)))

bbch_mpi_fgh_2014_mod$timepoint <- rep("only cultivars", nrow(bbch_mpi_fgh_2014_mod))
bbch_mpi_fgh_2014_mod$timepoint[bbch_mpi_fgh_2014_mod$date == "2014-05-12"] <- "early"
bbch_mpi_fgh_2014_mod$timepoint[bbch_mpi_fgh_2014_mod$date == "2014-05-26"] <- "late"
bbch_mpi_fgh_2014_mod$timepoint[bbch_mpi_fgh_2014_mod$date == "2014-05-27"] <- "late"
bbch_mpi_fgh_2014_mod$timepoint <- as.factor(bbch_mpi_fgh_2014_mod$timepoint)
table(bbch_mpi_fgh_2014_mod$timepoint)

# export datasets
write.table(bbch_trost_mpi, "data/bbch_trost_mpi.txt", sep="\t")
write.table(bbch_mpi_fgh_2014, "data/bbch_mpi_fgh_2014.txt", sep="\t")
```

## no data for late timepoint in mpitest2 experiment!

### TROST subsets
```{r TROST subsets}
# MPI TEST 1.2
bbch_mpi_test1_2 <- droplevels( subset(bbch_trost_mpi_mod, bbch_trost_mpi_mod$experiment_id == "48656") )
# reorder cultivar factor
levels(bbch_mpi_test1_2$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
bbch_mpi_test1_2$cultivar <- factor(bbch_mpi_test1_2$cultivar, levels = c("Alegria", "Milva", "Desiree", "Saturna"))

# ONLY LATE TIMEPOINT
bbch_mpi_test1_2_late <- droplevels (subset(bbch_mpi_test1_2, bbch_mpi_test1_2$timepoint=="late") )
# ONLY EARLY TIMEPOINT
bbch_mpi_test1_2_early <- droplevels( subset(bbch_mpi_test1_2, bbch_mpi_test1_2$timepoint=="early") )


# MPI TEST 2 (only early exists)
bbch_mpi_test2 <- droplevels( subset(bbch_trost_mpi_mod, bbch_trost_mpi_mod$experiment_id == "51790") )
# reorder cultivar factor
levels(bbch_mpi_test2$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
bbch_mpi_test2$cultivar <- factor(bbch_mpi_test2$cultivar, levels = c("Alegria", "Milva", "Desiree", "Saturna"))


# MPI FELD 2011
bbch_mpi_feld2011 <- droplevels( subset(bbch_trost_mpi_mod, bbch_trost_mpi_mod$experiment_id == "44443") )
# ONLY CHECK CULTIVARS
bbch_mpi_feld2011_check <- droplevels (subset(bbch_mpi_feld2011, 
                                              bbch_mpi_feld2011$cultivar %in% c("ALEGRIA", "MILVA", "DESIREE", "SATURNA")) )
# reorder cultivar factor
levels(bbch_mpi_feld2011_check$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
bbch_mpi_feld2011_check$cultivar <- factor(bbch_mpi_feld2011_check$cultivar, levels = c("Alegria", "Milva", "Desiree", "Saturna"))

# ONLY CHECK CULTIVARS + LATE
bbch_mpi_feld2011_check_late <- droplevels( subset(bbch_mpi_feld2011_check, 
                                                   bbch_mpi_feld2011_check$timepoint == "late") )
# ONLY LATE
bbch_mpi_feld2011_late <- droplevels( subset(bbch_mpi_feld2011, 
                                             bbch_mpi_feld2011$timepoint == "late") )


# MPI FELD 2012
bbch_mpi_feld2012 <- droplevels( subset(bbch_trost_mpi_mod, 
                                        bbch_trost_mpi_mod$experiment_id == "56726") )
# ONLY CHECK CULTIVARS 
bbch_mpi_feld2012_check <- droplevels( subset(bbch_mpi_feld2012, 
                                              bbch_mpi_feld2012$cultivar %in% c("ALEGRIA", "MILVA", "DESIREE", "SATURNA")) )
# reorder cultivar factor
levels(bbch_mpi_feld2012_check$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
bbch_mpi_feld2012_check$cultivar <- factor(bbch_mpi_feld2012_check$cultivar, levels = c("Alegria", "Milva", "Desiree", "Saturna"))
```

### VALDIS subsets
```{r VALDIS subsets}
# MPI FGH 2014
# ONLY LATE TIMEPOINT
bbch_mpi_fgh_2014_late <- droplevels (subset(bbch_mpi_fgh_2014_mod, bbch_mpi_fgh_2014_mod$timepoint=="late") )
# ONLY EARLY TIMEPOINT
bbch_mpi_fgh_2014_early <- droplevels (subset(bbch_mpi_fgh_2014_mod, bbch_mpi_fgh_2014_mod$timepoint=="early") )
```


### ggplot2 density plot
```{r ggplot2 density plot}
bbch_mpi_test1_2_late_part <- bbch_mpi_test1_2_late[,c("treatment","bbch")]
bbch_mpi_test1_2_early_part <- bbch_mpi_test1_2_early[,c("treatment","bbch")]

bbch_mpi_test2_part <- bbch_mpi_test2[,c("treatment","bbch")]

bbch_mpi_feld2011_check_late_part <- bbch_mpi_feld2011_check_late[,c("treatment","bbch")]
table(bbch_mpi_feld2011_check_late_part$bbch)

bbch_mpi_feld2011_late_part <- bbch_mpi_feld2011_late[,c("treatment","bbch")]
table(bbch_mpi_feld2011_late_part$bbch)

bbch_mpi_feld2012_check_late_part <- bbch_mpi_feld2012_check[,c("treatment","bbch")]
table(bbch_mpi_feld2012_check_late_part$bbch)


pdf("figures/bbch_mpi_test_1_2_late.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_test1_2_late_part) 
dev.off()

pdf("figures/bbch_mpi_test_1_2_early.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_test1_2_early_part) 
dev.off()

pdf("figures/bbch_mpi_test_2_early.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_test2_part) 
dev.off()

pdf("figures/bbch_mpi_feld_2011_check_late.pdf")
func_bbch_plots(bbch_mpi_feld2011_check_late_part) 
dev.off()

pdf("figures/bbch_mpi_feld_2012_check_late.pdf")
func_bbch_plots(bbch_mpi_feld2012_check_late_part) 
dev.off()
```


```{r for phd-thesis}
# for phd-thesis:
#pdf("~/work/Doktorarbeit/figures/bbch_mpitest1_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_test1_2_late_part)
#dev.off()

#pdf("~/work/Doktorarbeit/figures/bbch_mpi_field_2011_check_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_feld2011_check_late_part)
#dev.off()

#pdf("~/work/Doktorarbeit/figures/bbch_mpi_field_2011_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_feld2011_late_part)
#dev.off()

#pdf("~/work/Doktorarbeit/figures/bbch_mpi_field_2012_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_feld2012_check_late_part)
#dev.off()
```


```{r valdis BBCH plots}
bbch_mpi_fgh_2014_late_part <- bbch_mpi_fgh_2014_late[,c("treatment","bbch")]
bbch_mpi_fgh_2014_early_part <- bbch_mpi_fgh_2014_early[,c("treatment","bbch")]


pdf("figures/bbch_mpi_fgh_2014_mod_late.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_fgh_2014_late_part) 
dev.off()

pdf("figures/bbch_mpi_fgh_2014_mod_early.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_fgh_2014_early_part) 
dev.off()
```


### aggregate data --> median and sd
```{r aggregate data}
# aggregate by cultivar, experiment_id, treatment, timepoint --> median + sd
bbch_trost_mpi_median <- aggregate(bbch_trost_mpi_mod$bbch, 
                                   by=list(bbch_trost_mpi_mod$cultivar, 
                                           bbch_trost_mpi_mod$experiment_id, 
                                           bbch_trost_mpi_mod$treatment, 
                                           bbch_trost_mpi_mod$timepoint), 
                                   median)

bbch_trost_mpi_sd <- aggregate(bbch_trost_mpi_mod$bbch, 
                               by=list(bbch_trost_mpi_mod$cultivar, 
                                       bbch_trost_mpi_mod$experiment_id, 
                                       bbch_trost_mpi_mod$treatment, 
                                       bbch_trost_mpi_mod$timepoint), 
                               sd)

dim(bbch_trost_mpi_median)
# 228 5

colnames(bbch_trost_mpi_median) <- c("cultivar", "experiment_id", "treatment", "timepoint", "median") 
colnames(bbch_trost_mpi_sd) <- c("cultivar", "experiment_id", "treatment", "timepoint", "sd")

pander(head(bbch_trost_mpi_median))
pander(head(bbch_trost_mpi_sd))


# convert treatment column into 2 columns (1 for control, 1 for stress)
bbch_trost_mpi_median <- cast(bbch_trost_mpi_median, 
                                       cultivar + experiment_id + timepoint ~ treatment, 
                                       value="median")

bbch_trost_mpi_sd <- cast(bbch_trost_mpi_sd, 
                                     cultivar + experiment_id + timepoint ~ treatment, 
                                     value="sd")

colnames(bbch_trost_mpi_median)[5] <- "drought_stress"
colnames(bbch_trost_mpi_sd)[5] <- "drought_stress"

# combine median and sd
bbch_trost_mpi_agg <- cbind(bbch_trost_mpi_median, 
                            bbch_trost_mpi_sd$control,
                            bbch_trost_mpi_sd$drought_stress)

colnames(bbch_trost_mpi_agg)[6:7] <- c("sd_control", "sd_stress")
pander(head(bbch_trost_mpi_agg))

# only late timepoint, and only check cultivars
bbch_subset <- subset(bbch_trost_mpi_agg, 
           bbch_trost_mpi_agg$timepoint=="late" & 
             bbch_trost_mpi_agg$cultivar %in% c("ALEGRIA", "MILVA", "DESIREE", "SATURNA"))
dim(bbch_subset)
pander(bbch_subset)

write.table(bbch_subset, "output/bbch_late_check_cultivars.txt", sep="\t", quote=F, col.names=NA)
```


### rank transformation
```{r rank transformation}
# MPI greenhouse trial 1.2 (late)
dim(bbch_mpi_test1_2_late)
head(bbch_mpi_test1_2_late)
plot(density(bbch_mpi_test1_2_late$bbch, freq=F))
plot(density(rank(bbch_mpi_test1_2_late$bbch), freq=F))
shapiro.test(bbch_mpi_test1_2_late$bbch)
shapiro.test(rank(bbch_mpi_test1_2_late$bbch))

bbch_mpi_test1_2_late$bbch_rank <- rank(bbch_mpi_test1_2_late$bbch)

# MPI field 2011 (late)
dim(bbch_mpi_feld2011_check_late)
head(bbch_mpi_feld2011_check_late)
plot(density(bbch_mpi_feld2011_check_late$bbch, freq=F))
plot(density(rank(bbch_mpi_feld2011_check_late$bbch), freq=F))
shapiro.test(bbch_mpi_feld2011_check_late$bbch)
shapiro.test(rank(bbch_mpi_feld2011_check_late$bbch))

bbch_mpi_feld2011_check_late$bbch_rank <- rank(bbch_mpi_feld2011_check_late$bbch)

# MPI field 2012 (late)
dim(bbch_mpi_feld2012_check)
head(bbch_mpi_feld2012_check)
plot(density(bbch_mpi_feld2012_check$bbch, freq=F))
plot(density(rank(bbch_mpi_feld2012_check$bbch), freq=F))
shapiro.test(bbch_mpi_feld2012_check$bbch)
shapiro.test(rank(bbch_mpi_feld2012_check$bbch))

bbch_mpi_feld2012_check$bbch_rank <- rank(bbch_mpi_feld2012_check$bbch)
```


### ggplot2 density plot of ranks
```{r ggplot2 density plot of ranks}
bbch_mpi_test1_2_late_rank <- bbch_mpi_test1_2_late[,c("treatment","bbch_rank")]
colnames(bbch_mpi_test1_2_late_rank)[2] <- "bbch"

bbch_mpi_feld2011_check_late_rank <- bbch_mpi_feld2011_check_late[,c("treatment","bbch_rank")]
colnames(bbch_mpi_feld2011_check_late_rank)[2] <- "bbch"

bbch_mpi_feld2012_check_late_rank <- bbch_mpi_feld2012_check[,c("treatment","bbch_rank")]
colnames(bbch_mpi_feld2012_check_late_rank)[2] <- "bbch"


pdf("figures/bbch_mpi_test_1_2_late_rank.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_test1_2_late_rank, min=0, max=200, label_xaxis = "rank-transformed developmental stage (BBCH)") 
dev.off()

pdf("figures/bbch_mpi_feld_2011_check_late_rank.pdf")
func_bbch_plots(bbch_mpi_feld2011_check_late_rank, min=0, max=260, label_xaxis = "rank-transformed developmental stage (BBCH)") 
dev.off()

pdf("figures/bbch_mpi_feld_2012_check_late_rank.pdf")
func_bbch_plots(bbch_mpi_feld2012_check_late_rank, min=0, max=130, label_xaxis = "rank-transformed developmental stage (BBCH)") 
dev.off()
```

### ANOVA
```{r ANOVA of ranks}
bbch_mpi_test1_2_late_rank_anova <- func_anova_2fac_ia(bbch_mpi_test1_2_late, "bbch_rank", "cultivar", "treatment", 0.01)
  
bbch_mpi_feld2011_check_late_rank_anova <- func_anova_2fac_ia(bbch_mpi_feld2011_check_late, "bbch_rank", "cultivar", "treatment", 0.01)

bbch_mpi_feld2012_check_late_rank_anova <- func_anova_2fac_ia(bbch_mpi_feld2012_check, "bbch_rank", "cultivar", "treatment", 0.01)
```


### kruskal wallis test of ranks
```{r kruskal wallis test of ranks}
kruskal.test(bbch_mpi_test1_2_late$bbch_rank ~ bbch_mpi_test1_2_late$treatment)$p.value
kruskal.test(bbch_mpi_test1_2_late$bbch_rank ~ bbch_mpi_test1_2_late$treatment)$statistic
kruskal.test(bbch_mpi_test1_2_late$bbch_rank ~ bbch_mpi_test1_2_late$cultivar)$p.value
kruskal.test(bbch_mpi_test1_2_late$bbch_rank ~ bbch_mpi_test1_2_late$cultivar)$statistic

kruskal.test(bbch_mpi_feld2011_check_late$bbch_rank ~ bbch_mpi_feld2011_check_late$treatment)$p.value
kruskal.test(bbch_mpi_feld2011_check_late$bbch_rank ~ bbch_mpi_feld2011_check_late$treatment)$statistic
kruskal.test(bbch_mpi_feld2011_check_late$bbch_rank ~ bbch_mpi_feld2011_check_late$cultivar)$p.value
kruskal.test(bbch_mpi_feld2011_check_late$bbch_rank ~ bbch_mpi_feld2011_check_late$cultivar)$statistic

kruskal.test(bbch_mpi_feld2012_check$bbch_rank ~ bbch_mpi_feld2012_check$treatment)
kruskal.test(bbch_mpi_feld2012_check$bbch_rank ~ bbch_mpi_feld2012_check$cultivar)
```


### boxplots per treatment and cultivar of ranked BBCH
```{r boxplots per treatment and cultivar of ranked BBCH}
# MPI test 1.2 late

pdf("figures/bbch_mpi_test_1_2_late_rank_boxplot.pdf")
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(bbch_mpi_test1_2_late$bbch_rank ~
          bbch_mpi_test1_2_late$treatment * bbch_mpi_test1_2_late$cultivar,
        las = 2, names = names_treatment_cultivar, col = cols_treatment,
        main = "ranked BBCH")

boxplot(bbch_mpi_test1_2_late$bbch ~
          bbch_mpi_test1_2_late$treatment * bbch_mpi_test1_2_late$cultivar,
        las = 2, names = names_treatment_cultivar, col = cols_treatment,
        main = "original BBCH (no ranking)")
dev.off()


# MPI field 2011 (late)

pdf("figures/bbch_mpi_feld_2011_check_late_rank_boxplot.pdf")
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(bbch_mpi_feld2011_check_late$bbch_rank ~ 
          bbch_mpi_feld2011_check_late$treatment * bbch_mpi_feld2011_check_late$cultivar,
        las = 2, names = names_treatment_cultivar, col = cols_treatment,
        main = "ranked BBCH")

boxplot(bbch_mpi_feld2011_check_late$bbch ~ 
          bbch_mpi_feld2011_check_late$treatment * bbch_mpi_feld2011_check_late$cultivar,
        las = 2, names = names_treatment_cultivar, col = cols_treatment,
        main = "original BBCH (no ranking)")
dev.off()


# MPI field 2012 (late)

pdf("figures/bbch_mpi_feld_2012_check_late_rank_boxplot.pdf")
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(bbch_mpi_feld2012_check$bbch_rank ~ 
          bbch_mpi_feld2012_check$treatment * bbch_mpi_feld2012_check$cultivar,
        las = 2, names = names_treatment_cultivar, col = cols_treatment,
        main = "ranked BBCH")

boxplot(bbch_mpi_feld2012_check$bbch ~ 
          bbch_mpi_feld2012_check$treatment * bbch_mpi_feld2012_check$cultivar,
        las = 2, names = names_treatment_cultivar, col = cols_treatment,
        main = "original BBCH (no ranking)")
dev.off()
```


### aggregate ranked BBCH --> median and sd
#### aggregate by cultivar and treatment --> median + sd
```{r aggregate ranked BBCH}
# MPI greenhouse trial 1.2 (late)
bbch_mpi_test1_2_late_rank_agg <- ddply(bbch_mpi_test1_2_late, c("cultivar", "treatment"), summarise,
                                        N    = length(bbch_rank),
                                        mean = mean(bbch_rank),
                                        median = median(bbch_rank),
                                        sd   = sd(bbch_rank),
                                        se   = sd / sqrt(N) )

pander(bbch_mpi_test1_2_late_rank_agg)
write.table(bbch_mpi_test1_2_late_rank_agg, 
            "output/bbch_mpi_test1_2_late_rank_agg.txt",
            sep="\t", quote=F, col.names=NA)


# MPI field 2011 (late)
bbch_mpi_feld2011_check_late_rank_agg <- ddply(bbch_mpi_feld2011_check_late, c("cultivar", "treatment"), summarise,
                                        N    = length(bbch_rank),
                                        mean = mean(bbch_rank),
                                        median = median(bbch_rank),
                                        sd   = sd(bbch_rank),
                                        se   = sd / sqrt(N) )

pander(bbch_mpi_feld2011_check_late_rank_agg)
write.table(bbch_mpi_feld2011_check_late_rank_agg, 
            "output/bbch_mpi_feld2011_check_late_rank_agg.txt", 
            sep="\t", quote=F, col.names=NA)


# MPI field 2012 (late)
bbch_mpi_feld2012_check_late_rank_agg <- ddply(bbch_mpi_feld2012_check, c("cultivar", "treatment"), summarise,
                                        N    = length(bbch_rank),
                                        mean = mean(bbch_rank),
                                        median = median(bbch_rank),
                                        sd   = sd(bbch_rank),
                                        se   = sd / sqrt(N) )

pander(bbch_mpi_feld2012_check_late_rank_agg)
write.table(bbch_mpi_feld2012_check_late_rank_agg, 
            "output/bbch_mpi_feld2012_check_late_rank_agg.txt", 
            sep="\t", quote=F, col.names=NA)
```


### combine mean and sd
```{r combine mean and sd}
# G1:
mpi_test_1_2_median_sd <- func_combine_mean_sd(bbch_mpi_test1_2_late, variable_name = "bbch_rank", func = "median")

# F1:
mpi_feld2011_check_late_median_sd <- func_combine_mean_sd(bbch_mpi_feld2011_check_late, variable_name = "bbch_rank", func = "median")

# F3:
bbch_mpi_feld2012_check_median_sd <- func_combine_mean_sd(bbch_mpi_feld2012_check, variable_name = "bbch_rank", func = "median")
```


### compare control and stress by wilcoxon test
```{r compare control and stress by wilcoxon test}
wilcoxon_test_res <- c(func_wilcoxon_test_treatment(bbch_mpi_test1_2_late, variable_name = "bbch_rank"), #G1
                       func_wilcoxon_test_treatment(bbch_mpi_feld2011_check_late, variable_name = "bbch_rank")) #F1

```

**warning because of ties, for further information see**
[here](https://www.biostars.org/p/2524/)
[or here](http://r.789695.n4.nabble.com/What-are-ties-Wilcox-u-test-td857059.html)


### combine median and sd results of 2 experiments (for publication)
```{r combine median and sd results of 2 experiments}

bbch_check_median_sd <- rbind(mpi_test_1_2_median_sd, #G1
                              mpi_feld2011_check_late_median_sd) #F1

bbch_check_median_sd$trial <- c( rep("G1", 4), rep("F1", 4))

bbch_check_median_sd$percentage <- bbch_check_median_sd$median_drought_stress * 100 /
                                         bbch_check_median_sd$median_control

bbch_check_median_sd$p_value <- wilcoxon_test_res

pander(bbch_check_median_sd)

write.table(bbch_check_median_sd,
            "output/bbch_check_median_sd_pub.txt",
            sep="\t", quote=F, col.names = NA)
```

### save workspace
```{r save workspace}
save.image("BBCH.RData")
```

