BBCH data analysis for MPI trials
========================================================

## Set working directory  
```{r set working directory}
getwd()
setwd("D:/work/repos/trost_phenotypes")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

## Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(ggplot2)
library(reshape)
library(pander)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = 'D:/work/repos/trost_phenotypes/')

# load workspace
load("BBCH.RData")
```


## Source R functions
```{r source R functions, include=FALSE}
source("../libpurzel/func_get_bbch_data.R")
source("../libpurzel/func_modify_bbch_data.R")
source("../libpurzel/func_bbch_plots.R")
source("../libpurzel/colors.R")
```


## Execute Query from `D:\work\TROST\Database\2014-07-29-bbch_query.sql` to get information about BBCH and meta data
### BBCH data of four TROST experiments (MPI Test1.2, MPI Test2, MPI Feld 2011, MPI Feld 2012)
```{r get data from trost database}
# use old version of function
# bbch_trost_mpi_48656 <- func_get_bbch_data_old('48656')

bbch_trost_mpi <- func_get_bbch_data( experiment_id = c('48656', '51790', '44443', '56726') )

pander(head(bbch_trost_mpi))
dim(bbch_trost_mpi)
# 3842 11
```

## Execute Query from `D:\work\TROST\Database\2015-05-07-bbch_query_valdis.sql` to get information about BBCH and meta data
### Get BBCH data for VALDIS experiments at MPI FGH 2014
```{r get BBCH data for VALDIS experiments at MPI FGH 2014}
bbch_mpi_fgh_2014 <- func_get_bbch_data2( experiment_id = '67199')

pander(head(bbch_mpi_fgh_2014))
dim(bbch_mpi_fgh_2014)
# 2535 10
```


### Modification of BBCH dataset
```{r modification of BBCH dataset}
# TROST data
bbch_trost_mpi_mod <- func_modify_bbch_data(bbch_trost_mpi)
pander(table(bbch_trost_mpi_mod$experiment_name, bbch_trost_mpi_mod$timepoint))

# VALDIS data
bbch_mpi_fgh_2014_mod <- func_modify_bbch_data(bbch_mpi_fgh_2014)

# timepoints
table(bbch_mpi_fgh_2014_mod$date)
pander(head(table(bbch_mpi_fgh_2014_mod$line_id, bbch_mpi_fgh_2014_mod$date)))

bbch_mpi_fgh_2014_mod$timepoint <- rep("only cultivars", nrow(bbch_mpi_fgh_2014_mod))
bbch_mpi_fgh_2014_mod$timepoint[bbch_mpi_fgh_2014_mod$date == "2014-05-12"] <- "early"
bbch_mpi_fgh_2014_mod$timepoint[bbch_mpi_fgh_2014_mod$date == "2014-05-26"] <- "late"
bbch_mpi_fgh_2014_mod$timepoint[bbch_mpi_fgh_2014_mod$date == "2014-05-27"] <- "late"
bbch_mpi_fgh_2014_mod$timepoint <- as.factor(bbch_mpi_fgh_2014_mod$timepoint)
table(bbch_mpi_fgh_2014_mod$timepoint)

# export datasets
write.table(bbch_trost_mpi, "data/bbch_trost_mpi.txt", sep="\t")
write.table(bbch_mpi_fgh_2014, "data/bbch_mpi_fgh_2014.txt", sep="\t")
```

## no data for late timepoint in mpitest2 experiment!

### TROST subsets
```{r TROST subsets}
# MPI TEST 1.2
bbch_mpi_test1_2 <- droplevels( subset(bbch_trost_mpi_mod, bbch_trost_mpi_mod$experiment_id == "48656") )
# ONLY LATE TIMEPOINT
bbch_mpi_test1_2_late <- droplevels (subset(bbch_mpi_test1_2, bbch_mpi_test1_2$timepoint=="late") )
# ONLY EARLY TIMEPOINT
bbch_mpi_test1_2_early <- droplevels( subset(bbch_mpi_test1_2, bbch_mpi_test1_2$timepoint=="early") )

# MPI TEST 2 (only early exists)
bbch_mpi_test2 <- droplevels( subset(bbch_trost_mpi_mod, bbch_trost_mpi_mod$experiment_id == "51790") )
levels(bbch_mpi_test2$timepoint)

# MPI FELD 2011
bbch_mpi_feld2011 <- droplevels( subset(bbch_trost_mpi_mod, bbch_trost_mpi_mod$experiment_id == "44443") )
# ONLY CHECK CULTIVARS
bbch_mpi_feld2011_check <- droplevels (subset(bbch_mpi_feld2011, bbch_mpi_feld2011$cultivar %in% c("ALEGRIA", "MILVA", "DESIREE", "SATURNA")) )
levels(bbch_mpi_feld2011_check$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
# ONLY CHECK CULTIVARS + LATE
bbch_mpi_feld2011_check_late <- droplevels( subset(bbch_mpi_feld2011_check, bbch_mpi_feld2011_check$timepoint == "late") )
# ONLY LATE
bbch_mpi_feld2011_late <- droplevels( subset(bbch_mpi_feld2011, bbch_mpi_feld2011$timepoint == "late") )

# MPI FELD 2012
bbch_mpi_feld2012 <- droplevels( subset(bbch_trost_mpi_mod, bbch_trost_mpi_mod$experiment_id == "56726") )
# ONLY CHECK CULTIVARS 
bbch_mpi_feld2012_check <- droplevels( subset(bbch_mpi_feld2012, bbch_mpi_feld2012$cultivar %in% c("ALEGRIA", "MILVA", "DESIREE", "SATURNA")) )
levels(bbch_mpi_feld2012_check$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
```

### VALDIS subsets
```{r VALDIS subsets}
# MPI FGH 2014
# ONLY LATE TIMEPOINT
bbch_mpi_fgh_2014_late <- droplevels (subset(bbch_mpi_fgh_2014_mod, bbch_mpi_fgh_2014_mod$timepoint=="late") )
# ONLY EARLY TIMEPOINT
bbch_mpi_fgh_2014_early <- droplevels (subset(bbch_mpi_fgh_2014_mod, bbch_mpi_fgh_2014_mod$timepoint=="early") )
```


### ggplot2 density plot
```{r ggplot2 density plot}
bbch_mpi_test1_2_late_part <- bbch_mpi_test1_2_late[,c("treatment","bbch")]
bbch_mpi_test1_2_early_part <- bbch_mpi_test1_2_early[,c("treatment","bbch")]

bbch_mpi_test2_part <- bbch_mpi_test2[,c("treatment","bbch")]

bbch_mpi_feld2011_check_late_part <- bbch_mpi_feld2011_check_late[,c("treatment","bbch")]
table(bbch_mpi_feld2011_check_late_part$bbch)

bbch_mpi_feld2011_late_part <- bbch_mpi_feld2011_late[,c("treatment","bbch")]
table(bbch_mpi_feld2011_late_part$bbch)

bbch_mpi_feld2012_check_late_part <- bbch_mpi_feld2012_check[,c("treatment","bbch")]
table(bbch_mpi_feld2012_check_late_part$bbch)


pdf("figures/bbch_mpi_test_1_2_late.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_test1_2_late_part) 
dev.off()

pdf("figures/bbch_mpi_test_1_2_early.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_test1_2_early_part) 
dev.off()

pdf("figures/bbch_mpi_test_2_early.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_test2_part) 
dev.off()

pdf("figures/bbch_mpi_feld_2011_check_late.pdf")
func_bbch_plots(bbch_mpi_feld2011_check_late_part) 
dev.off()

pdf("figures/bbch_mpi_feld_2012_check_late.pdf")
func_bbch_plots(bbch_mpi_feld2012_check_late_part) 
dev.off()
```


```{r for phd-thesis}
# for phd-thesis:
#pdf("~/work/Doktorarbeit/figures/bbch_mpitest1_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_test1_2_late_part)
#dev.off()

#pdf("~/work/Doktorarbeit/figures/bbch_mpi_field_2011_check_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_feld2011_check_late_part)
#dev.off()

#pdf("~/work/Doktorarbeit/figures/bbch_mpi_field_2011_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_feld2011_late_part)
#dev.off()

#pdf("~/work/Doktorarbeit/figures/bbch_mpi_field_2012_late.pdf", width=6, height=6)
func_bbch_density_plot(bbch_mpi_feld2012_check_late_part)
#dev.off()
```


```{r valdis BBCH plots}
bbch_mpi_fgh_2014_late_part <- bbch_mpi_fgh_2014_late[,c("treatment","bbch")]
bbch_mpi_fgh_2014_early_part <- bbch_mpi_fgh_2014_early[,c("treatment","bbch")]


pdf("figures/bbch_mpi_fgh_2014_mod_late.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_fgh_2014_late_part) 
dev.off()

pdf("figures/bbch_mpi_fgh_2014_mod_early.pdf", width=6, height=6)
func_bbch_plots(bbch_mpi_fgh_2014_early_part) 
dev.off()
```


### aggregate data --> median and sd
```{r aggregate data}
# aggregate by cultivar, experiment_id, treatment, timepoint --> median + sd
bbch_trost_mpi_median <- aggregate(bbch_trost_mpi_mod$bbch, 
                                   by=list(bbch_trost_mpi_mod$cultivar, 
                                           bbch_trost_mpi_mod$experiment_id, 
                                           bbch_trost_mpi_mod$treatment, 
                                           bbch_trost_mpi_mod$timepoint), 
                                   median)

bbch_trost_mpi_sd <- aggregate(bbch_trost_mpi_mod$bbch, 
                               by=list(bbch_trost_mpi_mod$cultivar, 
                                       bbch_trost_mpi_mod$experiment_id, 
                                       bbch_trost_mpi_mod$treatment, 
                                       bbch_trost_mpi_mod$timepoint), 
                               sd)

dim(bbch_trost_mpi_median)
# 228 5

colnames(bbch_trost_mpi_median) <- c("cultivar", "experiment_id", "treatment", "timepoint", "median") 
colnames(bbch_trost_mpi_sd) <- c("cultivar", "experiment_id", "treatment", "timepoint", "sd")

pander(head(bbch_trost_mpi_median))
pander(head(bbch_trost_mpi_sd))


# convert treatment column into 2 columns (1 for control, 1 for stress)
bbch_trost_mpi_median <- cast(bbch_trost_mpi_median, 
                                       cultivar + experiment_id + timepoint ~ treatment, 
                                       value="median")

bbch_trost_mpi_sd <- cast(bbch_trost_mpi_sd, 
                                     cultivar + experiment_id + timepoint ~ treatment, 
                                     value="sd")

colnames(bbch_trost_mpi_median)[5] <- "drought_stress"
colnames(bbch_trost_mpi_sd)[5] <- "drought_stress"

# combine median and sd
bbch_trost_mpi_agg <- cbind(bbch_trost_mpi_median, 
                                       bbch_trost_mpi_sd$control,
                                       bbch_trost_mpi_sd$drought_stress)

colnames(bbch_trost_mpi_agg)[6:7] <- c("sd_control", "sd_stress")
pander(head(bbch_trost_mpi_agg))

# only late timepoint, and only check cultivars
bbch_subset <- subset(bbch_trost_mpi_agg, 
           bbch_trost_mpi_agg$timepoint=="late" & 
             bbch_trost_mpi_agg$cultivar %in% c("ALEGRIA", "MILVA", "DESIREE", "SATURNA"))
dim(bbch_subset)
pander(bbch_subset)

write.table(bbch_subset, "output/bbch_late_check_cultivars.txt", sep="\t", quote=F, col.names=NA)
```


### save workspace
```{r save workspace}
save.image("BBCH.RData")
```

