FW/DW data analysis for greenhouse and field trials
========================================================

### Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_phenotypes")
```


[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

### Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(ggplot2)
library(reshape)
library(reshape2)
library(pander)
library(plyr)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("fw_dw.RData")
```


### Source R functions
```{r source R functions, include=FALSE}
source("../libpurzel/colors.R")
source("../libpurzel/names.R")
source("../libpurzel/func_get_fw_dw_data.R")
source("../libpurzel/func_anova_v2.R")
source("../libpurzel/func_relevel_factors.R")
```

## Load Data
### Execute Query from D:\work\repos\database_scripts\plant_height_query-2014-07-07.sql to get information about plant height and meta data
```{r get MPI data from trost database}
# MPI trials
fw_dw_mpi_greenhouse <- func_get_fw_dw_mpi_greenhouse_data()
fw_dw_mpi_field <- func_get_fw_dw_mpi_field_data()

# JKI trials
fw_dw_jki_greenhouse <- func_get_fw_dw_jki_greenhouse_data()
fw_dw_jki_field <- func_get_fw_dw_jki_field_data()
```


### unique data
#### without plantID
```{r unique data}
# MPI trials
fw_dw_mpi_greenhouse_unique <- unique(fw_dw_mpi_greenhouse[ ,-1])
fw_dw_mpi_field_unique <- unique(fw_dw_mpi_field[ ,-1])

# JKI trials
fw_dw_jki_greenhouse_unique <- unique(fw_dw_jki_greenhouse[ ,-1])
fw_dw_jki_field_unique <- unique(fw_dw_jki_field[ ,-1])

# remove wrong entries
fw_dw_jki_field_correct <- fw_dw_jki_field_unique[-c(98,100) ,]
```


### combine data
```{r combine data}
fw_dw_all <- rbind(fw_dw_mpi_greenhouse_unique,
                   fw_dw_mpi_field_unique,
                   fw_dw_jki_greenhouse_unique,
                   fw_dw_jki_field_unique)

table(fw_dw_all$experiment_name)
table(fw_dw_all$experiment_name, fw_dw_all$description)
table(fw_dw_all$experiment_id, fw_dw_all$description)

write.table(fw_dw_all, "data/fw_dw_result_all.txt", sep="\t")
```


### subsets: only check cultivars at late or late/before
```{r subsets: only check cultivars at late or late/before}
# G1: MPI greenhouse trial 1.2
fw_dw_g1 <- droplevels(subset(fw_dw_all, 
                              fw_dw_all$experiment_id == "48656" & 
                                fw_dw_all$description == "late/before"))

# G2: MPI greenhouse trial 2
fw_dw_g2 <- droplevels(subset(fw_dw_all, 
                              fw_dw_all$experiment_id == "51790" & 
                                fw_dw_all$description == "late/before"))

# G3: JKI greenhouse trial 1
fw_dw_g3 <- droplevels(subset(fw_dw_all, 
                              fw_dw_all$experiment_id == "45985" & 
                                fw_dw_all$description == "late/before"))

# F1: MPI field trial 2011
fw_dw_f1 <- droplevels(subset(fw_dw_all, 
                              fw_dw_all$experiment_id == "44443" & 
                                fw_dw_all$description == "late" & 
                                fw_dw_all$cultivar %in% c("DESIREE", "ALEGRIA", "MILVA", "SATURNA")))

# F3: MPI field trial 2012
fw_dw_f3 <- droplevels(subset(fw_dw_all, 
                              fw_dw_all$experiment_id == "56726" & 
                                fw_dw_all$cultivar %in% c("DESIREE", "ALEGRIA", "MILVA", "SATURNA")))

# F4: JKI field trial 2012
fw_dw_f4 <- droplevels(subset(fw_dw_all, 
                              fw_dw_all$experiment_id == "56875" & 
                                fw_dw_all$cultivar %in% c("DESIREE", "ALEGRIA", "MILVA", "SATURNA")))
```

### relevel cultivar and treatment
```{r relevel cultivar and treatment}

fw_dw_g1 <- func_relevel_factors_2(fw_dw_g1)
fw_dw_g2 <- func_relevel_factors_2(fw_dw_g2)
fw_dw_g3 <- func_relevel_factors_2(fw_dw_g3)
fw_dw_f1 <- func_relevel_factors_2(fw_dw_f1)
fw_dw_f3 <- func_relevel_factors_2(fw_dw_f3)
fw_dw_f4 <- func_relevel_factors_2(fw_dw_f4)
```


### remove outlier
```{r remove outlier}
# remove outlier
hist(fw_dw_g1$FW_DW, col="grey", breaks=20)
hist(fw_dw_g2$FW_DW, col="grey", breaks=20)

hist(fw_dw_g3$FW_DW, col="grey", breaks=20)
fw_dw_g3 <- fw_dw_g3 [-which(fw_dw_g3$FW_DW > 20), ]

hist(fw_dw_f1$FW_DW, col="grey", breaks=20)
fw_dw_f1 <- fw_dw_f1 [-which(fw_dw_f1$FW_DW > 10), ]

hist(fw_dw_f3$FW_DW, col="grey", breaks=20)
fw_dw_f3 <- fw_dw_f3 [-which(fw_dw_f3$FW_DW > 10), ]
fw_dw_f3 <- fw_dw_f3 [-which(fw_dw_f3$FW_DW < 4), ]

hist(fw_dw_f4$FW_DW, col="grey", breaks=20)
```



## plots
### plot greenhouse trials
```{r plot greenhouse trials}
pdf("figures/fw_dw_greenhouse_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))

# G1: MPI greenhouse trial 1.2
boxplot(fw_dw_g1$FW_DW ~ fw_dw_g1$treatment, 
        col=cols_treatment, main="MPI test trial 1.2")
boxplot(fw_dw_g1$FW_DW ~ fw_dw_g1$treatment * 
          fw_dw_g1$cultivar, col=cols_treatment, las=2, 
        names=names_treatment_cultivar, ylab="FW/DW ratio", main="MPI test trial 1.2")

# G2: MPI greenhouse trial 2
boxplot(fw_dw_g2$FW_DW ~ fw_dw_g2$treatment, 
        col=cols_treatment, main="MPI test trial 2")
boxplot(fw_dw_g2$FW_DW ~ fw_dw_g2$treatment * 
          fw_dw_g2$cultivar, col=cols_treatment, las=2, 
        names=names_treatment_cultivar, ylab="FW/DW ratio", main="MPI test trial 2")

# G3: JKI greenhouse trial 1
boxplot(fw_dw_g3$FW_DW ~ fw_dw_g3$treatment, 
        col=cols_treatment, main="JKI test trial 1")
boxplot(fw_dw_g3$FW_DW ~ fw_dw_g3$treatment * 
          fw_dw_g3$cultivar, col=cols_treatment, las=2, 
        names=names_treatment_cultivar, ylab="FW/DW ratio", main="JKI test trial 1")

dev.off()
```


### plot field trials
```{r plot field trials}
pdf("figures/fw_dw_field_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))

# F1: MPI field trial 2011
boxplot(fw_dw_f1$FW_DW ~ fw_dw_f1$treatment, 
        col=cols_treatment, main="MPI field trial 2011")
boxplot(fw_dw_f1$FW_DW ~ fw_dw_f1$treatment * 
          fw_dw_f1$cultivar, col=cols_treatment, las=2, 
        names=names_treatment_cultivar, ylab="FW/DW ratio", main="MPI field trial 2011")

# F3: MPI field trial 2012
boxplot(fw_dw_f3$FW_DW ~ fw_dw_f3$treatment, 
        col=cols_treatment, main="MPI field trial 2012")
boxplot(fw_dw_f3$FW_DW ~ fw_dw_f3$treatment * 
          fw_dw_f3$cultivar, col=cols_treatment, las=2, 
        names=names_treatment_cultivar, ylab="FW/DW ratio", main="MPI field trial 2012")

# F4: JKI field trial 2012
boxplot(fw_dw_f4$FW_DW ~ fw_dw_f4$treatment, 
        col=cols_treatment, main="JKI field trial 2012")
boxplot(fw_dw_f4$FW_DW ~ fw_dw_f4$treatment * 
          fw_dw_f4$cultivar, col=cols_treatment, las=2, 
        names=names_treatment_cultivar, ylab="FW/DW ratio", main="JKI field trial 2012")

dev.off()
```



```{r aggregate with ddply}
# G1: MPI test 1.2 (late/before)
fw_dw_g1_agg <- ddply(fw_dw_g1, c("cultivar", "treatment"), summarise,
                      N    = sum(!is.na(FW_DW)),
                      mean = mean(FW_DW, na.rm=TRUE),
                      median = median(FW_DW, na.rm=TRUE),
                      sd   = sd(FW_DW, na.rm=TRUE),
                      se   = sd / sqrt(N) )
pander(fw_dw_g1_agg)


# G2: MPI test 2 (late/before)
fw_dw_g2_agg <- ddply(fw_dw_g2, c("cultivar", "treatment"), summarise,
                      N    = sum(!is.na(FW_DW)),
                      mean = mean(FW_DW, na.rm=TRUE),
                      median = median(FW_DW, na.rm=TRUE),
                      sd   = sd(FW_DW, na.rm=TRUE),
                      se   = sd / sqrt(N) )
pander(fw_dw_g2_agg)
```




```{r save workspace}
save.image("fw_dw.RData")
```

