Plant height data analysis for MPI greenhouse and field trials
========================================================

### Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_phenotypes")
```


[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

### Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(ggplot2)
library(reshape)
library(reshape2)
library(pander)
library(plyr)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("plant_height.RData")
```


### Source R functions
```{r source R functions, include=FALSE}
source("../functions/colors.R")
source("../functions/names.R")
source("../functions/func_get_plant_height_data.R")
source("../functions/func_anova_v2.R")
source("../functions/func_ttest.R")
source("../functions/func_aggregate_values.R")
```

## Load Data
### Execute Query from D:\work\repos\database_scripts\plant_height_query-2014-07-07.sql to get information about plant height and meta data
```{r get MPI data from trost database}
plant_height_query_result <- func_get_plant_height_data()

pander(head(plant_height_query_result))
dim(plant_height_query_result)
# 3808 11

write.table(plant_height_query_result, "data/plant_height_query_result.txt", sep="\t")
```


### load JKI test trial 1 data (45985)
plant height data for 7 dates, of which 15.07.2011 should be used 
date for late/before sampling: 12.07.2011, late/after: 15.07.2011
no BBCH data available!
```{r load JKI test trial 1 data}

plant_height_jki_test_1 <- read.table("data/JKI_GWH_Bonitur_2011_Original.txt", sep="\t", header=T)

# change level names
levels(plant_height_jki_test_1$treatment) <- c("control", "drought stress")
# change level order
plant_height_jki_test_1$cultivar <- factor(plant_height_jki_test_1$cultivar, check_names_ordered)


# only use cultivar, treatment and plant height columns (7 different dates)
plant_height_jki_test_1_part <- plant_height_jki_test_1[,c(3,4,8:14)]
# melt data for different timepoints
plant_height_jki_test_1_melt <- melt(plant_height_jki_test_1_part)

levels(plant_height_jki_test_1_melt$variable)

# subset data of sampling timepoint (15.07.2011)
plant_height_jki_test_1_late <- subset(plant_height_jki_test_1_melt, plant_height_jki_test_1_melt$variable == "height_15.07.2011")
```


### load JKI field trial 2012 data (56875)
BBCH and plant height data for 02.07.2012 (date for sampling)
plant height for 6 individual plants per ID
```{r load JKI field trial 2012 data}

plant_height_jki_field_2012 <- read.table("data/JKI_Feld_Bonitur_2012_Original.txt", sep="\t", header=T)

# change level names
levels(plant_height_jki_field_2012$treatment) <- c("control", "drought stress")

# only use cultivar, treatment and plant height columns (6 different plants)
plant_height_jki_field_2012_part <- plant_height_jki_field_2012[,c(3,5,8:13)]
# melt data for different plants
plant_height_jki_field_2012_melt <- melt(plant_height_jki_field_2012_part)

# only check cultivars
plant_height_jki_field_2012_check <- subset(plant_height_jki_field_2012, 
                                            plant_height_jki_field_2012$cultivar %in% check_names)
# change level order
plant_height_jki_field_2012_check$cultivar <- factor(plant_height_jki_field_2012_check$cultivar, check_names_ordered)

# only use cultivar, treatment and plant height columns (6 different plants)
plant_height_jki_field_2012_check_part <- plant_height_jki_field_2012_check[,c(3,5,8:13)]
# melt data for different plants
plant_height_jki_field_2012_check_melt <- melt(plant_height_jki_field_2012_check_part)
```


### MPI change factors
```{r MPI change factors}
# convert height to numeric values
plant_height_query_result$height <- as.numeric(plant_height_query_result$height)

# treatment
plant_height_query_result$treatment <- as.factor(plant_height_query_result$treatment)
levels(plant_height_query_result$treatment)
# rename factors
levels(plant_height_query_result$treatment) <- c("control", "drought stress")

# description of time point
plant_height_query_result$description <- as.factor(plant_height_query_result$description)
levels(plant_height_query_result$description)
# rename description
plant_height_query_result$description <- factor(plant_height_query_result$description, 
                                                levels=c("early","early/before", "early/after",
                                                         "late", "late/before", "late/after"))

# create timepoint factor
timepoint <- rep("early", nrow(plant_height_query_result))
timepoint [plant_height_query_result$description %in% c("late/before", "late/after", "late")] <- "late"
timepoint <- as.factor(timepoint)
levels(timepoint)

plant_height_query_result <- cbind(plant_height_query_result, timepoint)

pander(table(plant_height_query_result$experiment_name, plant_height_query_result$timepoint))

# cultivar
plant_height_query_result$cultivar <- as.factor(plant_height_query_result$cultivar)
levels(plant_height_query_result$cultivar) <- names_cultivars
```


### MPI greenhouse subsets
```{r MPI greenhouse subsets}
# MPI TEST 
plant_height_mpi_test1_2 <- subset(plant_height_query_result, plant_height_query_result$experiment_id == "48656")
plant_height_mpi_test2 <- subset(plant_height_query_result, plant_height_query_result$experiment_id == "51790")

plant_height_mpi_test1_2$cultivar <- droplevels(plant_height_mpi_test1_2$cultivar)
plant_height_mpi_test2$cultivar <- droplevels(plant_height_mpi_test2$cultivar)

# reorder cultivars
plant_height_mpi_test1_2$cultivar <- factor(plant_height_mpi_test1_2$cultivar, check_names_ordered)
plant_height_mpi_test2$cultivar <- factor(plant_height_mpi_test2$cultivar, check_names_ordered)

# ONLY LATE!
plant_height_mpi_test1_2_late <- subset(plant_height_mpi_test1_2, plant_height_mpi_test1_2$timepoint=="late")
plant_height_mpi_test2_late <- subset(plant_height_mpi_test2, plant_height_mpi_test2$timepoint=="late")

# early
plant_height_mpi_test1_2_early <- subset(plant_height_mpi_test1_2, plant_height_mpi_test1_2$timepoint=="early")
plant_height_mpi_test2_early <- subset(plant_height_mpi_test2, plant_height_mpi_test2$timepoint=="early")
```


### MPI field subsets
```{r MPI field subsets}
# MPI FIELD 2011
plant_height_mpi_field2011 <- subset(plant_height_query_result, plant_height_query_result$experiment_id == "44443")
# ONLY CHECK CULTIVARS
plant_height_mpi_field2011_check <- subset(plant_height_mpi_field2011, plant_height_mpi_field2011$cultivar %in% check_names)
# change level order
plant_height_mpi_field2011_check$cultivar <- factor(plant_height_mpi_field2011_check$cultivar, check_names_ordered)


# ONLY CHECK CULTIVARS + EARLY
plant_height_mpi_field2011_check_early <- subset(plant_height_mpi_field2011_check, plant_height_mpi_field2011_check$timepoint == "early")

# ONLY CHECK CULTIVARS + LATE
plant_height_mpi_field2011_check_late <- subset(plant_height_mpi_field2011_check, plant_height_mpi_field2011_check$timepoint == "late")


# ALL CULTIVARS + EARLY
plant_height_mpi_field2011_early <- subset(plant_height_mpi_field2011, plant_height_mpi_field2011$timepoint == "early")
# ALL CULTIVARS + LATE
plant_height_mpi_field2011_late <- subset(plant_height_mpi_field2011, plant_height_mpi_field2011$timepoint == "late")

####################################

# MPI FIELD 2012
plant_height_mpi_field2012 <- subset(plant_height_query_result, plant_height_query_result$experiment_id == "56726")
# ONLY CHECK CULTIVARS 
plant_height_mpi_field2012_check <- subset(plant_height_mpi_field2012, plant_height_mpi_field2012$cultivar %in% check_names)
# change level order
plant_height_mpi_field2012_check$cultivar <- factor(plant_height_mpi_field2012_check$cultivar, check_names_ordered)
```


## plots
### plot MPI test trials
```{r plot MPI test trials}
pdf("figures/plant_height_mpi_test_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))

# mpi test 1.2
boxplot(plant_height_mpi_test1_2$height ~ plant_height_mpi_test1_2$treatment, 
        col = cols_treatment, main = "MPI test trial 1.2")
boxplot(plant_height_mpi_test1_2$height ~ plant_height_mpi_test1_2$treatment * 
          plant_height_mpi_test1_2$timepoint, col = cols_treatment, las = 2, 
        names = names_treatment_sample_time, 
        ylab = "plant height (in cm)", main = "MPI test trial 1.2")

# mpi test 2
boxplot(plant_height_mpi_test2$height ~ plant_height_mpi_test2$treatment, 
        col = cols_treatment, main = "MPI test trial 2")
boxplot(plant_height_mpi_test2$height ~ plant_height_mpi_test2$treatment * 
          plant_height_mpi_test2$timepoint, col = cols_treatment, las = 2, 
        names = names_treatment_sample_time, 
        ylab = "plant height (in cm)", main = "MPI test trial 2")

dev.off()
```


### plot MPI test trials, per cultivar
```{r plot MPI test trials per cultivar}
pdf("figures/plant_height_mpi_test_trials_cultivar.pdf", width=12, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(plant_height_mpi_test1_2$height ~ plant_height_mpi_test1_2$treatment * 
          plant_height_mpi_test1_2$timepoint * plant_height_mpi_test1_2$cultivar, 
        col = cols_treatment_sample_time, las = 2, ylab = "plant height (in cm)", 
        main = "MPI test trial 1", names = names_treatment_sample_time_cultivar_reordered)

boxplot(plant_height_mpi_test2$height ~ plant_height_mpi_test2$treatment * 
          plant_height_mpi_test2$timepoint * plant_height_mpi_test2$cultivar, 
        col = cols_treatment_sample_time, las = 2, ylab = "plant height (in cm)", 
        main = "MPI test trial 2", names = names_treatment_sample_time_cultivar_reordered)

# ONLY LATE!
boxplot(plant_height_mpi_test1_2_late$height ~ plant_height_mpi_test1_2_late$treatment * 
          plant_height_mpi_test1_2_late$cultivar, col = cols_treatment, las = 2, 
        ylab = "plant height (in cm)", main = "MPI test trial 1 (only late)", 
        names = names_treatment_cultivar_reordered)
legend("topright", legend = levels(plant_height_mpi_test2_late$treatment), fill = cols_treatment)

boxplot(plant_height_mpi_test2_late$height ~ plant_height_mpi_test2_late$treatment * 
          plant_height_mpi_test2_late$cultivar, col = cols_treatment, las = 2, 
        ylab = "plant height (in cm)", main = "MPI test trial 2 (only late)", 
        names = names_treatment_cultivar_reordered)
legend("topright", legend = levels(plant_height_mpi_test2_late$treatment), fill = cols_treatment)
dev.off()
```


### plot MPI test trial for phd-thesis
```{r plot MPI test trial for phd-thesis}
# rename and reorder levels
# levels(plant_height_mpi_test1_2_late$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
# plant_height_mpi_test1_2_late$cultivar <- factor(plant_height_mpi_test1_2_late$cultivar, 
#                                                  levels=c("Alegria", "Milva", "Desiree", "Saturna"))

#pdf("~/work/Doktorarbeit/figures/plant_height_mpitest1_late.pdf", width=6, height=6)
par(mar=c(7.5, 4.5, 0.5, 0.5))
boxplot(plant_height_mpi_test1_2_late$height ~ plant_height_mpi_test1_2_late$treatment * 
          plant_height_mpi_test1_2_late$cultivar, col = cols_treatment, las = 2, 
        ylab = "plant height (in cm)", names = names_treatment_cultivar_reordered, 
        cex.lab=1.4, cex.axis=1.2, ylim=c(15,55))
legend("topright", levels(plant_height_mpi_test1_2_late$treatment), fill = cols_treatment, cex=1, bty="n")
#dev.off()
```


### plot JKI test trial
```{r plot JKI test trial}

pdf("figures/plant_height_jki_test_trial_1.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))

# only sampling timepoint (15.07.2011)
boxplot(plant_height_jki_test_1$height_15.07.2011 ~ plant_height_jki_test_1$treatment, 
        col = cols_treatment, main = "JKI test trial 1")

# all timepoints
boxplot(plant_height_jki_test_1_melt$value ~ plant_height_jki_test_1_melt$treatment * 
          plant_height_jki_test_1_melt$variable, 
        col = cols_treatment, main = "JKI test trial 1 (7 different timepoints)", las = 2, 
        names = names_treatment_sample_time_jki, ylab = "plant height (in cm)")

# per cultivar, only sampling timepoint
boxplot(plant_height_jki_test_1$height_15.07.2011 ~ plant_height_jki_test_1$treatment * 
          plant_height_jki_test_1$cultivar, 
        col = cols_treatment, las = 2, ylab = "plant height (in cm)", 
        main = "JKI test trial 1", names = names_treatment_cultivar)
legend("topright", legend = levels(plant_height_jki_test_1$treatment), fill = cols_treatment)

dev.off()
```


### plot MPI field trials
```{r plot MPI field trials}
pdf("figures/plant_height_mpi_field_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(plant_height_mpi_field2011_check$height ~ plant_height_mpi_field2011_check$treatment, 
        col = cols_treatment, main = "MPI field trial 2011")
boxplot(plant_height_mpi_field2011_check$height ~ plant_height_mpi_field2011_check$treatment *
          plant_height_mpi_field2011_check$timepoint, col = cols_treatment, main = "MPI field trial 2011")

boxplot(plant_height_mpi_field2012_check$height ~ plant_height_mpi_field2012_check$treatment, 
        col = cols_treatment, main = "MPI field trial 2012 (only late)")
# boxplot(plant_height_mpi_field2012_check$height ~ plant_height_mpi_field2012_check$treatment *
#           plant_height_mpi_field2012_check$timepoint, col = cols_treatment, main = "MPI field trial 2012")
dev.off()

# PER CULTIVAR

pdf("figures/plant_height_mpi_field_trials_cultivar.pdf", width=12, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(plant_height_mpi_field2011_check$height ~ plant_height_mpi_field2011_check$treatment*
          plant_height_mpi_field2011_check$timepoint * plant_height_mpi_field2011_check$cultivar, 
        col = cols_treatment_sample_time, main = "MPI field trial 2011", 
        names = names_treatment_sample_time_cultivar_reordered, las = 2)

boxplot(plant_height_mpi_field2011_check_late$height ~ plant_height_mpi_field2011_check_late$treatment*
          plant_height_mpi_field2011_check_late$cultivar, col = cols_treatment, 
        main = "MPI field trial 2011 (only late)", names = names_treatment_cultivar_reordered, las = 2)

boxplot(plant_height_mpi_field2012_check$height ~ plant_height_mpi_field2012_check$treatment* 
          plant_height_mpi_field2012_check$cultivar, col = cols_treatment, main = "MPI field trial 2012", 
        names = names_treatment_cultivar_reordered, las = 2)
dev.off()
```


### plot MPI field trial for phd-thesis
```{r plot MPI field trial for phd-thesis, fig.width=10}
range(plant_height_mpi_field2011_check_late$height)
range(plant_height_mpi_field2012_check$height)

# for phd-thesis:
#pdf("~/work/Doktorarbeit/figures/plant_height_mpi_field_2011_check_late.pdf", width=6, height=6)
par(mar=c(7.5, 4.5, 0.5, 0.5))
boxplot(plant_height_mpi_field2011_check_late$height ~ plant_height_mpi_field2011_check_late$treatment * 
          plant_height_mpi_field2011_check_late$cultivar, col = cols_treatment, las = 2, 
        ylab = "plant height (in cm)", names = names_treatment_cultivar_reordered, cex.lab=1.4, cex.axis=1.2)
legend("topleft", levels(plant_height_mpi_field2011_check_late$treatment), fill = cols_treatment, cex=1, bty="n")
#dev.off()

#pdf("~/work/Doktorarbeit/figures/plant_height_mpi_field_2012_check_late.pdf", width=6, height=6)
par(mar=c(7.5, 4.5, 0.5, 0.5))
boxplot(plant_height_mpi_field2012_check$height ~ plant_height_mpi_field2012_check$treatment * 
          plant_height_mpi_field2012_check$cultivar, col = cols_treatment, las = 2, 
        ylab = "plant height (in cm)", names = names_treatment_cultivar_reordered, cex.lab=1.4, cex.axis=1.2)
legend("topleft", levels(plant_height_mpi_field2012_check$treatment), fill = cols_treatment, cex=1, bty="n")
#dev.off()


### ALL Cultivars (only late)
#pdf("~/work/Doktorarbeit/figures/plant_height_mpi_field_2011_late.pdf", width=10, height=6)
par(mar=c(7, 4.5, 0.5, 0.5))
boxplot(plant_height_mpi_field2011_late$height ~ plant_height_mpi_field2011_late$treatment * plant_height_mpi_field2011_late$cultivar, col = cols_treatment, las = 2, ylab = "plant height (in cm)", names = names_cultivar_all, cex.lab=1.4, cex.axis=1.2)
legend("topleft", levels(plant_height_mpi_field2011_late$treatment), fill = cols_treatment, cex=1, bty="n")
#dev.off()
```


### plot JKI field trial
```{r plot JKI field trial}
pdf("figures/plant_height_jki_field_trial_2012.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))

# all cultivars
boxplot(plant_height_jki_field_2012_melt$value ~ plant_height_jki_field_2012_melt$treatment, 
        col = cols_treatment, main = "JKI field trial 2012 (all cultivars)")

# only cultivars
boxplot(plant_height_jki_field_2012_check_melt$value ~ plant_height_jki_field_2012_check_melt$treatment, 
        col = cols_treatment, main = "JKI field trial 2012 (only check cultivars)")

# per check cultivar
boxplot(plant_height_jki_field_2012_check_melt$value ~ plant_height_jki_field_2012_check_melt$treatment * 
          plant_height_jki_field_2012_check_melt$cultivar, 
        col = cols_treatment, las = 2, ylab = "plant height (in cm)", 
        main = "JKI field trial 2012", names = names_treatment_cultivar_reordered)
legend("topright", legend = levels(plant_height_jki_test_1$treatment), fill = cols_treatment)

dev.off()
```


### aggregate data --> mean and sd
```{r aggregate data}
# aggregate by cultivar, experiment_id, treatment, timepoint --> mean + sd
plant_height_query_result_mean <- aggregate(plant_height_query_result$height, 
                                            by=list(plant_height_query_result$cultivar, 
                                                    plant_height_query_result$experiment_id, 
                                                    plant_height_query_result$treatment, 
                                                    plant_height_query_result$timepoint), 
                                            mean)

plant_height_query_result_sd <- aggregate(plant_height_query_result$height, 
                                          by=list(plant_height_query_result$cultivar, 
                                                  plant_height_query_result$experiment_id, 
                                                  plant_height_query_result$treatment, 
                                                  plant_height_query_result$timepoint), 
                                          sd)

dim(plant_height_query_result_mean)
# 236 5

colnames(plant_height_query_result_mean) <- c("cultivar", "experiment_id", "treatment", "timepoint", "mean") 
colnames(plant_height_query_result_sd) <- c("cultivar", "experiment_id", "treatment", "timepoint", "sd")

pander(head(plant_height_query_result_mean))
pander(head(plant_height_query_result_sd))

# convert treatment column into 2 columns (1 for control, 1 for stress)
plant_height_query_result_mean <- cast(plant_height_query_result_mean, 
                                       cultivar + experiment_id + timepoint ~ treatment, 
                                       value="mean")

plant_height_query_result_sd <- cast(plant_height_query_result_sd, 
                                     cultivar + experiment_id + timepoint ~ treatment, 
                                     value="sd")

colnames(plant_height_query_result_mean)[5] <- "drought_stress"
colnames(plant_height_query_result_sd)[5] <- "drought_stress"

# combine mean and sd
plant_height_query_result_agg <- cbind(plant_height_query_result_mean, 
                                       plant_height_query_result_sd$control,
                                       plant_height_query_result_sd$drought_stress)

colnames(plant_height_query_result_agg)[6:7] <- c("sd_control", "sd_stress")
pander(head(plant_height_query_result_agg))

# only late timepoint, and only check cultivars
plant_height_subset <- subset(plant_height_query_result_agg, 
           plant_height_query_result_agg$timepoint=="late" & 
             plant_height_query_result_agg$cultivar %in% check_names)
dim(plant_height_subset)
pander(plant_height_subset)

write.table(plant_height_subset, "output/plant_height_late_check_cultivars.txt", sep="\t", quote=F, col.names = NA)
```

#### aggregate subsets
```{r aggregate subsets}
######################
# aggregated subsets #
######################

plant_height_mpi_test_agg <- subset(plant_height_query_result_agg, 
                                    plant_height_query_result_agg$experiment_id %in% c("48656", "51790"))

plant_height_mpi_field_agg <- subset(plant_height_query_result_agg, 
                                    plant_height_query_result_agg$experiment_id %in% c("44443", "56726"))

plant_height_mpi_test1_2_agg <- subset(plant_height_query_result_agg, 
                                       plant_height_query_result_agg$experiment_id == "48656")
plant_height_mpi_test2_agg <- subset(plant_height_query_result_agg, 
                                     plant_height_query_result_agg$experiment_id == "51790")
dim(plant_height_mpi_test2_agg)
# 16 5
pander(plant_height_mpi_test2_agg)

plant_height_mpi_field2011_agg <- subset(plant_height_query_result_agg, 
                                        plant_height_query_result_agg$experiment_id == "44443")
plant_height_mpi_field2012_agg <- subset(plant_height_query_result_agg, 
                                        plant_height_query_result_agg$experiment_id == "51790")
```


### ANOVA
```{r ANOVA}
# MPI test 1.2/2 EARLY
plant_height_mpi_test1_2_early_anova <- func_anova_2fac_ia(plant_height_mpi_test1_2_early, 
                                                           "height", "cultivar", "treatment", 0.01)
plant_height_mpi_test2_early_anova <- func_anova_2fac_ia(plant_height_mpi_test2_early, 
                                                         "height", "cultivar", "treatment", 0.01)

# MPI test 1.2/2 LATE
plant_height_mpi_test1_2_late_anova <- func_anova_2fac_ia(plant_height_mpi_test1_2_late, 
                                                          "height", "cultivar", "treatment", 0.01)
plant_height_mpi_test2_late_anova <- func_anova_2fac_ia(plant_height_mpi_test2_late, 
                                                        "height", "cultivar", "treatment", 0.01)

# JKI test 1 LATE
plant_height_jki_test1_late_anova <- func_anova_2fac_ia(plant_height_jki_test_1_late, 
                                                        "value", "cultivar", "treatment", 0.01)

# MPI field 2011/2012 late (only check cultivars)
plant_height_mpi_field2011_check_late_anova <- func_anova_2fac_ia(plant_height_mpi_field2011_check_late, 
                                                                 "height", "cultivar", "treatment", 0.01)
plant_height_mpi_field2012_check_anova <- func_anova_2fac_ia(plant_height_mpi_field2012_check, 
                                                            "height", "cultivar", "treatment", 0.01)

# MPI field 2011 early (only check cultivars)
plant_height_mpi_field2011_check_early_anova <- func_anova_2fac_ia(plant_height_mpi_field2011_check_early, 
                                                                  "height", "cultivar", "treatment", 0.01)

# JKI field 2012 LATE
plant_height_jki_field_2012_check_anova <- func_anova_2fac_ia(plant_height_jki_field_2012_check_melt, 
                                                        "value", "cultivar", "treatment", 0.01)
```


### combine ANOVA results
```{r combine ANOVA results}
# combine anova results of 4 experiments (only LATE)
plant_height_anova <- func_combine_anova_res(plant_height_mpi_test1_2_late_anova, 
                       plant_height_mpi_test2_late_anova, 
                       plant_height_mpi_field2011_check_late_anova, 
                       plant_height_mpi_field2012_check_anova,
                       "f_value","p_value", "treatment", "cultivar")

rownames(plant_height_anova) <- c("mpitest1_2", "mpitest2", "mpif_field_2011", "mpi_field_2012")
write.table(plant_height_anova, "output/plant_height_anova.txt", sep="\t", quote=F, row.names = T, col.names = NA)

####################################################################################

# combine anova results of 7 experiments (EARLY AND LATE)
plant_height_anova7 <- func_combine_anova_res7(plant_height_mpi_test1_2_early_anova, 
                                             plant_height_mpi_test1_2_late_anova,
                                             plant_height_mpi_test2_early_anova,
                                             plant_height_mpi_test2_late_anova, 
                                             plant_height_mpi_field2011_check_early_anova, 
                                             plant_height_mpi_field2011_check_late_anova,
                                             plant_height_mpi_field2012_check_anova,
                                             "f_value","p_value", "treatment", "cultivar")

rownames(plant_height_anova7) <- c("MPI-MP greenhouse trial 1 early", "MPI-MP greenhouse trial 1 late", 
                                   "MPI-MP greenhouse trial 2 early", "MPI-MP greenhouse trial 2 late",
                                   "MPI-MP field trial 2011 early", "MPI-MP field trial 2011 late", 
                                   "MPI-MP field trial 2012 late")
write.table(plant_height_anova7, "output/plant_height_anova7.txt", sep="\t", quote=F, row.names = T, col.names = NA)

####################################################################################

# combine anova results of 6 experiments (for publication)
plant_height_anova6 <- func_combine_anova_res6(plant_height_mpi_test1_2_late_anova,
                                             plant_height_mpi_test2_late_anova, 
                                             plant_height_jki_test1_late_anova,
                                             plant_height_mpi_field2011_check_late_anova,
                                             plant_height_mpi_field2012_check_anova,
                                             plant_height_jki_field_2012_check_anova,
                                             "f_value","p_value", "treatment", "cultivar")

rownames(plant_height_anova6) <- c("G1", "G2", "G3", "F1", "F3", "F4")
write.table(plant_height_anova6, "output/plant_height_anova_pub.txt", sep="\t", quote=F, row.names = T, col.names = NA)


#### with interactions!

plant_height_anova6_ia <- func_combine_anova_res6_ia(plant_height_mpi_test1_2_late_anova,
                                             plant_height_mpi_test2_late_anova, 
                                             plant_height_jki_test1_late_anova,
                                             plant_height_mpi_field2011_check_late_anova,
                                             plant_height_mpi_field2012_check_anova,
                                             plant_height_jki_field_2012_check_anova,
                                             "f_value","p_value", "treatment", "cultivar")
rownames(plant_height_anova6_ia) <- c("G1", "G2", "G3", "F1", "F3", "F4")
write.table(plant_height_anova6_ia, "output/plant_height_anova_pub_with_ia.txt", sep="\t", quote=F, row.names = T, col.names = NA)
```


### aggregate with ddply
```{r aggregate with ddply}
# MPI test 1.2 (late)
plant_height_mpi_test1_2_late_agg <- ddply(plant_height_mpi_test1_2_late, c("cultivar", "treatment"), summarise,
                             N    = sum(!is.na(height)),
                             mean = mean(height, na.rm=TRUE),
                             median = median(height, na.rm=TRUE),
                             sd   = sd(height, na.rm=TRUE),
                             se   = sd / sqrt(N) )
pander(plant_height_mpi_test1_2_late_agg)

# MPI test 2 (late)
plant_height_mpi_test2_late_agg <- ddply(plant_height_mpi_test2_late, c("cultivar", "treatment"), summarise,
                             N    = sum(!is.na(height)),
                             mean = mean(height, na.rm=TRUE),
                             median = median(height, na.rm=TRUE),
                             sd   = sd(height, na.rm=TRUE),
                             se   = sd / sqrt(N) )
pander(plant_height_mpi_test2_late_agg)

# JKI test 1 (late)
plant_height_jki_test_1_late_agg <- ddply(plant_height_jki_test_1_late, c("cultivar", "treatment"), summarise,
                             N    = sum(!is.na(value)),
                             mean = mean(value, na.rm=TRUE),
                             median = median(value, na.rm=TRUE),
                             sd   = sd(value, na.rm=TRUE),
                             se   = sd / sqrt(N) )
pander(plant_height_jki_test_1_late_agg)

# MPI field 2011 (late)
plant_height_mpi_field2011_check_late_agg <- ddply(plant_height_mpi_field2011_check_late, c("cultivar", "treatment"), summarise,
                             N    = sum(!is.na(height)),
                             mean = mean(height, na.rm=TRUE),
                             median = median(height, na.rm=TRUE),
                             sd   = sd(height, na.rm=TRUE),
                             se   = sd / sqrt(N) )
pander(plant_height_mpi_field2011_check_late_agg)

# MPI field 2012 (late)
plant_height_mpi_field2012_check_late_agg <- ddply(plant_height_mpi_field2012_check, c("cultivar", "treatment"), summarise,
                             N    = sum(!is.na(height)),
                             mean = mean(height, na.rm=TRUE),
                             median = median(height, na.rm=TRUE),
                             sd   = sd(height, na.rm=TRUE),
                             se   = sd / sqrt(N) )
pander(plant_height_mpi_field2012_check_late_agg)

# JKI field 2012 (late)
plant_height_jki_field_2012_check_agg <- ddply(plant_height_jki_field_2012_check_melt, c("cultivar", "treatment"), summarise,
                             N    = sum(!is.na(value)),
                             mean = mean(value, na.rm=TRUE),
                             median = median(value, na.rm=TRUE),
                             sd   = sd(value, na.rm=TRUE),
                             se   = sd / sqrt(N) )
pander(plant_height_jki_field_2012_check_agg)
```


### combine aggregated data
```{r combine aggregated data}
plant_height_late_check_agg <- rbind(plant_height_mpi_test1_2_late_agg, #G1
                                     plant_height_mpi_test2_late_agg, #G2
                                     plant_height_jki_test_1_late_agg, #G3
                                     plant_height_mpi_field2011_check_late_agg, #F1
                                     plant_height_mpi_field2012_check_late_agg, #F3
                                     plant_height_jki_field_2012_check_agg) #F4

plant_height_late_check_agg$trial <- c( rep("G1", 8), rep("G2", 8), rep("G3", 8),
                                        rep("F1", 8), rep("F3", 8), rep("F4", 8))

write.table(plant_height_late_check_agg,
            "output/plant_height_late_check_agg.txt",
            sep="\t", quote=F, col.names = NA)
```


### combine mean and sd
```{r combine mean and sd}
plant_height_mpi_test1_2_late_mean_sd <- func_combine_mean_sd(plant_height_mpi_test1_2_late, 
                                                              variable_name = "height")

plant_height_mpi_test2_late_mean_sd <- func_combine_mean_sd(plant_height_mpi_test2_late, 
                                                              variable_name = "height")

plant_height_jki_test_1_late_mean_sd <- func_combine_mean_sd(plant_height_jki_test_1_late, 
                                                              variable_name = "value")

plant_height_mpi_field2011_check_late_mean_sd <- func_combine_mean_sd(plant_height_mpi_field2011_check_late, 
                                                              variable_name = "height")

plant_height_mpi_field2012_check_late_mean_sd <- func_combine_mean_sd(plant_height_mpi_field2012_check, 
                                                              variable_name = "height")

plant_height_jki_field_2012_check_mean_sd <- func_combine_mean_sd(plant_height_jki_field_2012_check_melt, 
                                                              variable_name = "value")
```


### t-test: compare control and stress
```{r t-test}
plant_height_ttest_res <- c(func_ttest_treatment2(plant_height_mpi_test1_2_late, variable_name = "height"), #G1
                            func_ttest_treatment2(plant_height_mpi_test2_late, variable_name = "height"), #G2
                            func_ttest_treatment2(plant_height_jki_test_1_late, variable_name = "value"), #G3
                            func_ttest_treatment2(plant_height_mpi_field2011_check_late, variable_name = "height"), #F1
                            func_ttest_treatment2(plant_height_mpi_field2012_check, variable_name = "height"), #F3
                            func_ttest_treatment2(plant_height_jki_field_2012_check_melt, variable_name = "value")) #F4
```


### combine mean and sd results of all experiments
```{r combine mean and sd results of all experiments}

plant_height_late_check_mean_sd <- rbind(plant_height_mpi_test1_2_late_mean_sd, #G1
                                     plant_height_mpi_test2_late_mean_sd, #G2
                                     plant_height_jki_test_1_late_mean_sd, #G3
                                     plant_height_mpi_field2011_check_late_mean_sd, #F1
                                     plant_height_mpi_field2012_check_late_mean_sd, #F3
                                     plant_height_jki_field_2012_check_mean_sd) #F4

plant_height_late_check_mean_sd$trial <- c( rep("G1", 4), rep("G2", 4), rep("G3", 4),
                                        rep("F1", 4), rep("F3", 4), rep("F4", 4))

plant_height_late_check_mean_sd$percent_loss <- 100- (plant_height_late_check_mean_sd$mean_drought_stress * 100 /
                                                plant_height_late_check_mean_sd$mean_control)


plant_height_late_check_mean_sd$p_value <- plant_height_ttest_res

pander(head(plant_height_late_check_mean_sd))


write.table(plant_height_late_check_mean_sd,
            "output/plant_height_late_check_mean_sd.txt",
            sep="\t", quote=F, col.names = NA)
```

### `plant_height_late_check_mean_sd` is used for supplementary table S2


```{r save workspace}
save.image("plant_height.RData")
```

