RWC data analysis for MPI greenhouse trials
========================================================

```{r set up}
setwd("~/work/TROST/Auswertung/RWC/scripts/")
load("rwc.RData")

source("colors.R")
```


### Execute Query from H:\work\TROST\Database\2014-07-07-plant_height_query.sql to get information about RWC and meta data
```{r get data from trost database}
source("../functions/func_get_rwc_data.R")
rwc_query_result <- func_get_rwc_data()

head(rwc_query_result)
dim(rwc_query_result)
# 3212 13

write.table(rwc_query_result, "../data/rwc_query_result.txt", sep="\t")
```


### factors
```{r factors}
levels(rwc_query_result$treatment)
levels(rwc_query_result$treatment) <- c("control", "drought stress")

levels(rwc_query_result$description)
# change order of description
rwc_query_result$description <- factor(rwc_query_result$description, levels=c("early/before", "early/after", "late/before", "late/after"))
```


### correct and remove outlier
```{r correct and remove outlier}
rwc_query_result[which(rwc_query_result$RWC>2),]
which(rwc_query_result$RWC>2)
# 910

rwc_query_result$FW[910] <- 1.415
rwc_query_result$RWC[910] <- 0.807393
rwc_query_result[910,]

#################

# remove outlier
rwc_query_result[which(rwc_query_result$RWC>1.1),]
which(rwc_query_result$RWC>1.1)
rwc_query_result <- rwc_query_result[-c(803, 903, 2494),]

write.table(rwc_query_result, "../data/rwc_query_result.txt", sep="\t")
```

### subsets
```{r subsets}
rwc_test_trials <- subset(rwc_query_result, rwc_query_result$experiment_id %in% c("48656", "51790"))

# rename and reorder cultivar factor
rwc_test_trials$cultivar <- droplevels(rwc_test_trials$cultivar)
levels(rwc_test_trials$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
rwc_test_trials$cultivar <- factor(rwc_test_trials$cultivar, levels=c("Alegria", "Milva", "Desiree", "Saturna"))

rwc_pruef_trials <- subset(rwc_query_result, rwc_query_result$experiment_id %in% c("56575", "58243", "60319"))

rwc_test1 <- subset(rwc_test_trials, rwc_test_trials$experiment_id == "48656")
rwc_test2 <- subset(rwc_test_trials, rwc_test_trials$experiment_id == "51790")

rwc_pruef1 <- subset(rwc_query_result, rwc_query_result$experiment_id == "56575")
rwc_pruef2 <- subset(rwc_query_result, rwc_query_result$experiment_id == "58243")
rwc_pruef3 <- subset(rwc_query_result, rwc_query_result$experiment_id == "60319")
```


```{r subsets --> timepoints}
rwc_test1_early_before <- subset(rwc_test1, rwc_test1$description == "early/before")
rwc_test2_early_before <- subset(rwc_test2, rwc_test2$description == "early/before")

rwc_test1_early_after <- subset(rwc_test1, rwc_test1$description == "early/after")
rwc_test2_early_after <- subset(rwc_test2, rwc_test2$description == "early/after")


rwc_test1_late_before <- subset(rwc_test1, rwc_test1$description == "late/before")
rwc_test2_late_before <- subset(rwc_test2, rwc_test2$description == "late/before")

rwc_test1_late_after <- subset(rwc_test1, rwc_test1$description == "late/after")
rwc_test2_late_after <- subset(rwc_test2, rwc_test2$description == "late/after")

# only desiree
rwc_test1_desiree <- subset(rwc_test1, rwc_test1$cultivar == "Desiree")
rwc_test1_desiree$description
```



### names for plot labels
```{r names for plot labels}
names_treatment_sample_time <- c("control \n early/before", "drought stress \n early/before", "control \n early/after", "drought stress \n early/after", "control \n late/before", "drought stress \n late/before", "control \n late/after", "drought stress \n late/after")

names_treatment_cultivar_reordered <- c("control \n Alegria", 
                              "drought stress \n Alegria", 
                              "control \n Milva", 
                              "drought stress \n Milva", 
                              "control \n Desiree", 
                              "drought stress \n Desiree",
                              "control \n Saturna", 
                              "drought stress \n Saturna")
```


### plots test trials
```{r plots test trials}
pdf("../figures/rwc_test_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))

boxplot(rwc_test_trials$RWC~rwc_test_trials$treatment, col=c("blue", "red"), ylab=" relative water content (RWC)", las=2)
boxplot(rwc_test_trials$RWC~rwc_test_trials$treatment * rwc_test_trials$description, col=c("blue", "red"), las=2, names=names_treatment_sample_time, ylab=" relative water content (RWC)")

# mpi test 1.2
boxplot(rwc_test1$RWC~rwc_test1$treatment, col=c("blue", "red"), main="MPI test trial 1.2")
boxplot(rwc_test1$RWC~rwc_test1$treatment * rwc_test1$description, col=c("blue", "red"), las=2, names=names_treatment_sample_time, ylab=" relative water content (RWC)", main="MPI test trial 1.2")
boxplot(rwc_test1$RWC~rwc_test1$treatment * rwc_test1$cultivar, col=c("blue", "red"), las=2, ylab=" relative water content (RWC)", main="MPI test trial 1")
boxplot(rwc_test1_late_before$RWC~rwc_test1_late_before$treatment * rwc_test1_late_before$cultivar, col=c("blue", "red"), las=2, ylab=" relative water content (RWC)", main="MPI test trial 1 (late/before)", names=names_treatment_cultivar_reordered)

boxplot(rwc_test1_desiree$RWC ~ rwc_test1_desiree$treatment * rwc_test1_desiree$description, col=c("blue", "red"), las=2, ylab="relative water content (RWC)", names=names_treatment_sample_time, cex.lab=1.4, cex.axis=1.2, main="MPI test trial 1.2 (Desiree)")

# mpi test 2
boxplot(rwc_test2$RWC~rwc_test2$treatment, col=c("blue", "red"), main="MPI test trial 2")
boxplot(rwc_test2$RWC~rwc_test2$treatment * rwc_test2$description, col=c("blue", "red"), las=2, names=names_treatment_sample_time, ylab=" relative water content (RWC)", main="MPI test trial 2")
boxplot(rwc_test2$RWC~rwc_test2$treatment * rwc_test2$cultivar, col=c("blue", "red"), las=2, ylab=" relative water content (RWC)", main="MPI test trial 2")
boxplot(rwc_test2_late_before$RWC~rwc_test2_late_before$treatment * rwc_test2_late_before$cultivar, col=c("blue", "red"), las=2, ylab=" relative water content (RWC)", main="MPI test trial 2 (late/before)", names=names_treatment_cultivar_reordered)
dev.off()


# for phd-thesis:
pdf("~/work/Doktorarbeit/figures/rwc_mpitest1_late_before.pdf", width=6, height=6)
par(mar=c(7.5, 4.5, 0.5, 0.5))
boxplot(rwc_test1_late_before$RWC~rwc_test1_late_before$treatment * rwc_test1_late_before$cultivar, col=cols_treatment, las=2, ylab=" relative water content (RWC)", names=names_treatment_cultivar_reordered, cex.lab=1.4, cex.axis=1.2)
legend("topright", levels(rwc_test1_late_before$treatment), fill=cols_treatment, cex=1, bty="n")
dev.off()
```

### plots pruef trials
```{r plots pruef trials}
pdf("../figures/rwc_pruef_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(rwc_pruef_trials$RWC~rwc_pruef_trials$treatment, col=c("blue", "red"), ylab=" relative water content (RWC)", las=2)
boxplot(rwc_pruef_trials$RWC~rwc_pruef_trials$treatment * rwc_pruef_trials$description, col=c("blue", "red"), las=2, names=names_treatment_sample_time, ylab=" relative water content (RWC)")

boxplot(rwc_pruef1$RWC~rwc_pruef1$treatment, col=c("blue", "red"), main="pruef1")
boxplot(rwc_pruef1$RWC~rwc_pruef1$treatment * rwc_pruef1$description, col=c("blue", "red"), main="pruef1")

boxplot(rwc_pruef2$RWC~rwc_pruef2$treatment, col=c("blue", "red"), main="pruef2")
boxplot(rwc_pruef2$RWC~rwc_pruef2$treatment * rwc_pruef2$description, col=c("blue", "red"), main="pruef2")

boxplot(rwc_pruef3$RWC~rwc_pruef3$treatment, col=c("blue", "red"), main="pruef3")
boxplot(rwc_pruef3$RWC~rwc_pruef3$treatment * rwc_pruef3$description, col=c("blue", "red"), main="pruef3")
dev.off()
```

### aggregate data per trial, timepoint, cultivar and treatment
```{r aggregate data per trial, timepoint, cultivar and treatment}
# aggregate by experiment_id, cultivar, treatment, description
rwc_query_result_agg <- aggregate(rwc_query_result[,10:13], by=list(rwc_query_result$experiment_id, rwc_query_result$cultivar, rwc_query_result$treatment, rwc_query_result$description), mean)

dim(rwc_query_result_agg)
head(rwc_query_result_agg)
colnames(rwc_query_result_agg)[1:4] <- c("experiment_id", "cultivar", "treatment", "description") 


######################
# aggregated subsets #
######################

rwc_test_trials_agg <- subset(rwc_query_result_agg, rwc_query_result_agg$experiment_id %in% c("48656", "51790"))

rwc_pruef_trials_agg <- subset(rwc_query_result_agg, rwc_query_result_agg$experiment_id %in% c("56575", "58243", "60319"))

rwc_test1_agg <- subset(rwc_query_result_agg, rwc_query_result_agg$experiment_id == "48656")
rwc_test2_agg <- subset(rwc_query_result_agg, rwc_query_result_agg$experiment_id == "51790")

rwc_pruef1_agg <- subset(rwc_query_result_agg, rwc_query_result_agg$experiment_id == "56575")
rwc_pruef2_agg <- subset(rwc_query_result_agg, rwc_query_result_agg$experiment_id == "58243")
rwc_pruef3_agg <- subset(rwc_query_result_agg, rwc_query_result_agg$experiment_id == "60319")
```


### aggregate data per trial, timepoint and treatment
```{r aggregate data per trial, timepoint and treatment}
# aggregate by experiment_id, cultivar, treatment, description
rwc_query_result_agg2 <- aggregate(rwc_query_result[,10:13], by=list(rwc_query_result$treatment, rwc_query_result$description, rwc_query_result$experiment_id), mean)

dim(rwc_query_result_agg2)
head(rwc_query_result_agg2)
colnames(rwc_query_result_agg2)[1:3] <- c("treatment", "description", "experiment_id") 


######################
# aggregated subsets #
######################

rwc_test_trials_agg2 <- subset(rwc_query_result_agg2, rwc_query_result_agg2$experiment_id %in% c("48656", "51790"))

rwc_pruef_trials_agg2 <- subset(rwc_query_result_agg2, rwc_query_result_agg2$experiment_id %in% c("56575", "58243", "60319"))

rwc_test1_agg2 <- subset(rwc_query_result_agg2, rwc_query_result_agg2$experiment_id == "48656")
rwc_test2_agg2 <- subset(rwc_query_result_agg2, rwc_query_result_agg2$experiment_id == "51790")

rwc_pruef1_agg2 <- subset(rwc_query_result_agg2, rwc_query_result_agg2$experiment_id == "56575")
rwc_pruef2_agg2 <- subset(rwc_query_result_agg2, rwc_query_result_agg2$experiment_id == "58243")
rwc_pruef3_agg2 <- subset(rwc_query_result_agg2, rwc_query_result_agg2$experiment_id == "60319")
```


### ANOVA 
```{r ANOVA}
source("../functions/func_anova.R")

rwc_test1_early_before_anova <- func_anova_2fac_ia(rwc_test1_early_before, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_early_before_anova <- func_anova_2fac_ia(rwc_test2_early_before, "RWC", "cultivar", "treatment", 0.01)

rwc_test1_early_after_anova <- func_anova_2fac_ia(rwc_test1_early_after, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_early_after_anova <- func_anova_2fac_ia(rwc_test2_early_after, "RWC", "cultivar", "treatment", 0.01)

rwc_test1_late_before_anova <- func_anova_2fac_ia(rwc_test1_late_before, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_late_before_anova <- func_anova_2fac_ia(rwc_test2_late_before, "RWC", "cultivar", "treatment", 0.01)

rwc_test1_late_after_anova <- func_anova_2fac_ia(rwc_test1_late_after, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_late_after_anova <- func_anova_2fac_ia(rwc_test2_late_after, "RWC", "cultivar", "treatment", 0.01)


# combine anova results of 4 experiments
rwc_anova <- func_combine_anova_res(rwc_test1_early_before_anova, 
                                             rwc_test1_early_after_anova,
                                             rwc_test1_late_before_anova,
                                             rwc_test1_late_after_anova,
                                             rwc_test2_early_before_anova, 
                                             rwc_test2_early_after_anova,
                                             rwc_test2_late_before_anova,
                                             rwc_test2_late_after_anova,
                                             "f_value","p_value", "treatment", "cultivar")

rownames(rwc_anova) <- c("mpitest1_early/before", "mpitest1_early/after", "mpitest1_late/before", "mpitest1_late/after",
                         "mpitest2_early/before", "mpitest2_early/after", "mpitest2_late/before", "mpitest2_late/after")
write.table(rwc_anova, "../output/rwc_anova.txt", sep="\t", quote=F, row.names=T, col.names=NA)
```



```{r save workspace}
save.image("rwc.RData")
```

