RWC data analysis for MPI greenhouse trials
========================================================

### set working directory
```{r set working directory}
# setwd("D:/work/repos/trost_phenotypes")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

### Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(ggplot2)
library(reshape)
library(reshape2)
library(pander)
library(plyr)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = 'D:/work/repos/trost_phenotypes/')

# load workspace
#load("rwc.RData")
```


### source R functions
```{r source R functions, include=FALSE}
source("../libpurzel/func_get_rwc_data.R")
source("../libpurzel/colors.R")
source("../libpurzel/names.R")
source("../libpurzel/func_anova_v2.R")
```


### Execute Query from D:\work\repos\database_scripts\rwc_query-2014-06-24.sql to get information about RWC and meta data
```{r get data from trost database}
rwc_query_result <- func_get_rwc_data()

head(rwc_query_result)
dim(rwc_query_result)
# 3212 13

write.table(rwc_query_result, "data/rwc_query_result.txt", sep="\t")
```


### factors
```{r factors}
levels(rwc_query_result$treatment)
# rename treatment levels
levels(rwc_query_result$treatment) <- c("control", "drought stress")

levels(rwc_query_result$description)
# change order of description
rwc_query_result$description <- factor(rwc_query_result$description, levels=c("early/before", "early/after", "late/before", "late/after"))
```


### correct and remove outlier
```{r correct and remove outlier}
rwc_query_result[which(rwc_query_result$RWC>2),]
which(rwc_query_result$RWC>2)
# 910

rwc_query_result$FW[910] <- 1.415
rwc_query_result$RWC[910] <- 0.807393
rwc_query_result[910,]

#################

# remove outlier
rwc_query_result[which(rwc_query_result$RWC>1.1),]
which(rwc_query_result$RWC>1.1)
rwc_query_result <- rwc_query_result[-c(803, 903, 2494),]

write.table(rwc_query_result, "data/rwc_query_result_corrected.txt", sep="\t")
```


### subsets
```{r subsets}
rwc_test_trials <- subset(rwc_query_result, 
                          rwc_query_result$experiment_id %in% c("48656", "51790"))

# rename and reorder cultivar factor
rwc_test_trials$cultivar <- droplevels(rwc_test_trials$cultivar)
levels(rwc_test_trials$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
rwc_test_trials$cultivar <- factor(rwc_test_trials$cultivar, 
                                   levels=c("Alegria", "Milva", "Desiree", "Saturna"))

rwc_pruef_trials <- subset(rwc_query_result, 
                           rwc_query_result$experiment_id %in% c("56575", "58243", "60319"))

rwc_test1 <- subset(rwc_test_trials, rwc_test_trials$experiment_id == "48656")
rwc_test2 <- subset(rwc_test_trials, rwc_test_trials$experiment_id == "51790")

rwc_pruef1 <- subset(rwc_query_result, rwc_query_result$experiment_id == "56575")
rwc_pruef2 <- subset(rwc_query_result, rwc_query_result$experiment_id == "58243")
rwc_pruef3 <- subset(rwc_query_result, rwc_query_result$experiment_id == "60319")
```


### subsets --> timepoints
```{r subsets --> timepoints}
rwc_test1_early_before <- subset(rwc_test1, rwc_test1$description == "early/before")
rwc_test2_early_before <- subset(rwc_test2, rwc_test2$description == "early/before")

rwc_test1_early_after <- subset(rwc_test1, rwc_test1$description == "early/after")
rwc_test2_early_after <- subset(rwc_test2, rwc_test2$description == "early/after")


rwc_test1_late_before <- subset(rwc_test1, rwc_test1$description == "late/before")
rwc_test2_late_before <- subset(rwc_test2, rwc_test2$description == "late/before")

rwc_test1_late_after <- subset(rwc_test1, rwc_test1$description == "late/after")
rwc_test2_late_after <- subset(rwc_test2, rwc_test2$description == "late/after")

# only desiree
rwc_test1_desiree <- subset(rwc_test1, rwc_test1$cultivar == "Desiree")
rwc_test1_desiree$description
```


### plots test trials
```{r plots test trials}
pdf("figures/rwc_test_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))

boxplot(rwc_test_trials$RWC ~ rwc_test_trials$treatment, 
        col = cols_treatment, ylab = "relative water content (RWC)", las = 2)

boxplot(rwc_test_trials$RWC ~ rwc_test_trials$treatment * 
          rwc_test_trials$description, col = cols_treatment, las = 2, 
        names = names_treatment_sample_time_2, ylab = "relative water content (RWC)")


# mpi test 1.2
boxplot(rwc_test1$RWC ~ rwc_test1$treatment, 
        col = cols_treatment, main = "MPI test trial 1.2")

boxplot(rwc_test1$RWC ~ rwc_test1$treatment * rwc_test1$description, 
        col = cols_treatment, las = 2, names = names_treatment_sample_time_2, 
        ylab = "relative water content (RWC)", main = "MPI test trial 1.2")

boxplot(rwc_test1$RWC ~ rwc_test1$treatment * rwc_test1$cultivar, 
        col = cols_treatment, las = 2, ylab = "relative water content (RWC)", 
        main = "MPI test trial 1")

boxplot(rwc_test1_late_before$RWC ~ rwc_test1_late_before$treatment * 
          rwc_test1_late_before$cultivar, 
        col = cols_treatment, las = 2, 
        ylab = "relative water content (RWC)", 
        main = "MPI test trial 1 (late/before)", 
        names = names_treatment_cultivar_reordered)

boxplot(rwc_test1_desiree$RWC ~ rwc_test1_desiree$treatment * 
          rwc_test1_desiree$description, 
        col = cols_treatment, las = 2, ylab = "relative water content (RWC)", 
        names = names_treatment_sample_time_2, cex.lab=1.4, cex.axis=1.2, 
        main = "MPI test trial 1.2 (Desiree)")

# mpi test 2
boxplot(rwc_test2$RWC ~ rwc_test2$treatment, 
        col = cols_treatment, main = "MPI test trial 2")
boxplot(rwc_test2$RWC ~ rwc_test2$treatment * 
          rwc_test2$description, 
        col = cols_treatment, las = 2, 
        names = names_treatment_sample_time_2, 
        ylab = "relative water content (RWC)", 
        main = "MPI test trial 2")

boxplot(rwc_test2$RWC ~ rwc_test2$treatment * 
          rwc_test2$cultivar, 
        col = cols_treatment, las = 2, 
        ylab = "relative water content (RWC)", 
        main = "MPI test trial 2")

boxplot(rwc_test2_late_before$RWC ~ rwc_test2_late_before$treatment * 
          rwc_test2_late_before$cultivar, 
        col = cols_treatment, las = 2, 
        ylab = "relative water content (RWC)", 
        main = "MPI test trial 2 (late/before)", 
        names = names_treatment_cultivar_reordered)
dev.off()
```


### plot for phd thesis
```{r plot for phd thesis}
# for phd-thesis:
#pdf("~/work/Doktorarbeit/figures/rwc_mpitest1_late_before.pdf", width=6, height=6)
par(mar=c(7.5, 4.5, 0.5, 0.5))
boxplot(rwc_test1_late_before$RWC ~ rwc_test1_late_before$treatment * 
          rwc_test1_late_before$cultivar, 
        col = cols_treatment, las = 2, 
        ylab = "relative water content (RWC)", 
        names = names_treatment_cultivar_reordered, 
        cex.lab = 1.4, cex.axis = 1.2)
legend("topright", 
       levels(rwc_test1_late_before$treatment), 
       fill = cols_treatment, 
       cex = 1, bty = "n")
#dev.off()
```


### plots pruef trials
```{r plots pruef trials}
pdf("figures/rwc_pruef_trials.pdf", width=7, height=8)
par(mar=c(7, 4.1, 4.1, 2.1))
boxplot(rwc_pruef_trials$RWC ~ rwc_pruef_trials$treatment, 
        col = cols_treatment, 
        ylab = "relative water content (RWC)", las = 2)

boxplot(rwc_pruef_trials$RWC ~ rwc_pruef_trials$treatment * 
          rwc_pruef_trials$description, 
        col = cols_treatment, las = 2, 
        names = names_treatment_sample_time_2, 
        ylab = "relative water content (RWC)")

boxplot(rwc_pruef1$RWC ~ rwc_pruef1$treatment, 
        col = cols_treatment, main = "pruef1")

boxplot(rwc_pruef1$RWC ~ rwc_pruef1$treatment * 
          rwc_pruef1$description, 
        col = cols_treatment, main = "pruef1")

boxplot(rwc_pruef2$RWC ~ rwc_pruef2$treatment, 
        col = cols_treatment, main = "pruef2")

boxplot(rwc_pruef2$RWC ~ rwc_pruef2$treatment * 
          rwc_pruef2$description, 
        col = cols_treatment, main = "pruef2")


boxplot(rwc_pruef3$RWC ~ rwc_pruef3$treatment, 
        col = cols_treatment, main = "pruef3")

boxplot(rwc_pruef3$RWC ~ rwc_pruef3$treatment * 
          rwc_pruef3$description, 
        col = cols_treatment, main = "pruef3")
dev.off()
```

### aggregate data per trial, timepoint, cultivar and treatment
```{r aggregate data per trial- timepoint- cultivar and treatment}
# aggregate by experiment_id, cultivar, treatment, description
rwc_query_result_agg <- aggregate(rwc_query_result[,10:13], 
                                  by = list(rwc_query_result$experiment_id, 
                                            rwc_query_result$cultivar,
                                            rwc_query_result$treatment, 
                                            rwc_query_result$description), 
                                  mean)

dim(rwc_query_result_agg)
head(rwc_query_result_agg)
colnames(rwc_query_result_agg)[1:4] <- c("experiment_id", "cultivar", "treatment", "description") 


######################
# aggregated subsets #
######################

rwc_test_trials_agg <- subset(rwc_query_result_agg, 
                              rwc_query_result_agg$experiment_id %in% c("48656", "51790"))

rwc_pruef_trials_agg <- subset(rwc_query_result_agg, 
                               rwc_query_result_agg$experiment_id %in% c("56575", "58243", "60319"))

rwc_test1_agg <- subset(rwc_query_result_agg, 
                        rwc_query_result_agg$experiment_id == "48656")

rwc_test2_agg <- subset(rwc_query_result_agg, 
                        rwc_query_result_agg$experiment_id == "51790")

rwc_pruef1_agg <- subset(rwc_query_result_agg, 
                         rwc_query_result_agg$experiment_id == "56575")

rwc_pruef2_agg <- subset(rwc_query_result_agg, 
                         rwc_query_result_agg$experiment_id == "58243")

rwc_pruef3_agg <- subset(rwc_query_result_agg, 
                         rwc_query_result_agg$experiment_id == "60319")
```


### aggregate data per trial, timepoint and treatment
```{r aggregate data per trial- timepoint and treatment}
# aggregate by experiment_id, cultivar, treatment, description
rwc_query_result_agg2 <- aggregate(rwc_query_result[,10:13], 
                                   by = list(rwc_query_result$treatment, 
                                             rwc_query_result$description, 
                                             rwc_query_result$experiment_id), 
                                   mean)

dim(rwc_query_result_agg2)
head(rwc_query_result_agg2)
colnames(rwc_query_result_agg2)[1:3] <- c("treatment", "description", "experiment_id") 


######################
# aggregated subsets #
######################

rwc_test_trials_agg2 <- subset(rwc_query_result_agg2, 
                               rwc_query_result_agg2$experiment_id %in% c("48656", "51790"))

rwc_pruef_trials_agg2 <- subset(rwc_query_result_agg2, 
                                rwc_query_result_agg2$experiment_id %in% c("56575", "58243", "60319"))

rwc_test1_agg2 <- subset(rwc_query_result_agg2, 
                         rwc_query_result_agg2$experiment_id == "48656")

rwc_test2_agg2 <- subset(rwc_query_result_agg2, 
                         rwc_query_result_agg2$experiment_id == "51790")

rwc_pruef1_agg2 <- subset(rwc_query_result_agg2, 
                          rwc_query_result_agg2$experiment_id == "56575")

rwc_pruef2_agg2 <- subset(rwc_query_result_agg2, 
                          rwc_query_result_agg2$experiment_id == "58243")

rwc_pruef3_agg2 <- subset(rwc_query_result_agg2, 
                          rwc_query_result_agg2$experiment_id == "60319")
```


### ANOVA 
```{r ANOVA}
rwc_test1_early_before_anova <- func_anova_2fac_ia(rwc_test1_early_before, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_early_before_anova <- func_anova_2fac_ia(rwc_test2_early_before, "RWC", "cultivar", "treatment", 0.01)

rwc_test1_early_after_anova <- func_anova_2fac_ia(rwc_test1_early_after, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_early_after_anova <- func_anova_2fac_ia(rwc_test2_early_after, "RWC", "cultivar", "treatment", 0.01)

rwc_test1_late_before_anova <- func_anova_2fac_ia(rwc_test1_late_before, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_late_before_anova <- func_anova_2fac_ia(rwc_test2_late_before, "RWC", "cultivar", "treatment", 0.01)

rwc_test1_late_after_anova <- func_anova_2fac_ia(rwc_test1_late_after, "RWC", "cultivar", "treatment", 0.01)
rwc_test2_late_after_anova <- func_anova_2fac_ia(rwc_test2_late_after, "RWC", "cultivar", "treatment", 0.01)


# combine anova results of 4 timepoints of MPI test 1.2
rwc_test1_anova <- func_combine_anova_res(rwc_test1_early_before_anova, 
                                          rwc_test1_early_after_anova,
                                          rwc_test1_late_before_anova,
                                          rwc_test1_late_after_anova,
                                          "f_value","p_value", "treatment", "cultivar")

rwc_test2_anova <- func_combine_anova_res(rwc_test2_early_before_anova, 
                                          rwc_test2_early_after_anova,
                                          rwc_test2_late_before_anova,
                                          rwc_test2_late_after_anova,
                                          "f_value","p_value", "treatment", "cultivar")

rownames(rwc_test1_anova) <- c("mpitest1_early/before", "mpitest1_early/after", 
                               "mpitest1_late/before", "mpitest1_late/after")

rownames(rwc_test2_anova) <- c("mpitest2_early/before", "mpitest2_early/after", 
                               "mpitest2_late/before", "mpitest2_late/after")

write.table(rwc_test1_anova, "output/rwc_mpi_test1_2_anova.txt", sep="\t", quote=F, row.names = T, col.names = NA)
write.table(rwc_test2_anova, "output/rwc_mpi_test2_anova.txt", sep="\t", quote=F, row.names = T, col.names = NA)
```



```{r save workspace}
save.image("rwc.RData")
```

