---
title: "Tolerance classification and Phenotyper Metadata for TROST Prediction Model"
author: "Heike Sprenger"
date: "Monday, March 10, 2014"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---


# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_phenotypes/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(ggplot2)
library(plyr)
library(reshape)
library(RMySQL)
library(yaml)
library(gplots)
library(pander)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("tolerance_classification.RData")
```


# Establish connection to phenotyper DB
```{r establish connection to phenotyper DB}
login <- yaml.load_file("../functions/login.yaml")
phenotyper <- dbConnect(MySQL(), user=login$user, password=login$passwd, dbname=login$db, host=login$host) 

query_send <- dbSendQuery(phenotyper, 'select *  from species')
dbFetch(query_send)
```


# Metadata for Model
## Subspecies table --> get maturity group
```{r subspecies query}
subspecies_query <- paste("SELECT * FROM subspecies")

subspecies_query_result <- dbFetch( dbSendQuery(phenotyper, subspecies_query) )
head(subspecies_query_result)
write.table(subspecies_query_result, "output/subspecies_query_result.txt", sep="\t")

# select only the 34 cultivars which were used for TROST
subspecies_query_result_part <- subspecies_query_result[5:38,]
# summary of maturity groups
table(subspecies_query_result_part$reifegrclass)
table(subspecies_query_result_part$reifegruppe) # msp-sp and sp are both class 5 (2 = fr, 3 = mfr)

# change colname
colnames(subspecies_query_result_part)[1] <- "subspecies_id"
```


## Plant Height
Execute plant height query from ``D:\work\TROST\Database\prediction_model_query.sql`` to get information about the plant height of the 3 MPI Pruef trials:
- Pruef1: 56575
- Pruef2: 58243
- Pruef3: 60319

```{r plant height query}
plant_height_query <- paste("SELECT pl.id, pl.name, pl.subspecies_id, pl.culture_id, p.date, p.time, p.entity_id, p.value_id, p.number
                              FROM plants AS pl
                              LEFT JOIN phenotype_plants AS pp ON pp.plant_id = pl.id
                              LEFT JOIN phenotypes AS p ON p.id = pp.phenotype_id
                              WHERE p.value_id = 53 AND pl.culture_id in(56575, 58243, 60319);")

plant_height_query_result <- fetch( dbSendQuery(phenotyper, plant_height_query) , n=-1)
head(plant_height_query_result)

plant_height_query_result$number <- as.integer(plant_height_query_result$number)
#write.table(plant_height_query_result, "output/plant_height_query_result.txt", sep="\t")
```


## Treatment: Control
Execute to get plant with the treatment ID 169 (= CONTROL) of the 3 MPI Pruef trials:
- Pruef1: 56575
- Pruef2: 58243
- Pruef3: 60319

```{r treatment query}
treatment_control_query <- paste("SELECT pl.id, pl.name, pl.subspecies_id, pl.culture_id, p.date, p.time, p.entity_id, p.value_id, p.number
                              FROM plants AS pl
                              LEFT JOIN phenotype_plants AS pp ON pp.plant_id = pl.id
                              LEFT JOIN phenotypes AS p ON p.id = pp.phenotype_id
                              WHERE p.value_id = 169 AND pl.culture_id in(56575, 58243, 60319);")

treatment_control_query_result <- dbFetch( dbSendQuery(phenotyper, treatment_control_query) , n=-1)
head(treatment_control_query_result)
#write.table(treatment_control_query_result, "output/treatment_control_query_result.txt", sep="\t")
```


## Merge Plant Height with Control treated plants
```{r merge plant height with control}
plant_height_control <- join(treatment_control_query_result, plant_height_query_result, by="id")
head(plant_height_control)
dim(plant_height_control)

# only selected columns
plant_height_control <- plant_height_control[,c(1:4,8,13:17)]
head(plant_height_control)

# number of values per date: for Pruef3 (60319), late there are TWO phenotyping dates
table(plant_height_control$date)
table(format(plant_height_control$date, format="%b %Y"))

# sort plant height control by date 
plant_height_control <- plant_height_control[order(plant_height_control$date, plant_height_control$subspecies_id),]

# add column for timepoint (early/late)
time_point <- c(rep("early", 204), rep("late", 204), rep("early", 204), rep("late", 204), rep("early", 204), rep("late", 203))
plant_height_control <- cbind(plant_height_control, time_point)
```


## Aggregate plant height (use median) 
1. by culture, timepoint and cultivar
2. then by culture and timepoint to get median over all cultivars
```{r aggregate plant height by other factors}
# 1
# aggregate by culture_id, time_point and cultivar
plant_height_control_agg <- aggregate(plant_height_control$number, 
                                      by=list(plant_height_control$subspecies_id, 
                                              plant_height_control$time_point, 
                                              plant_height_control$culture_id), 
                                      median, na.rm=TRUE)

colnames(plant_height_control_agg) <- c("subspecies_id", "time_point", "culture_id", "number")
head(plant_height_control_agg)
dim(plant_height_control_agg)

# 2
# aggregate by culture_id and time_point (AFTER aggregate by culture_id, time_point and cultivar)!
plant_height_control_agg_1 <- aggregate(plant_height_control_agg$number,
                                        by=list(plant_height_control_agg$time_point,
                                                plant_height_control_agg$culture_id),
                                        median)
colnames(plant_height_control_agg_1) <- c("time_point","culture_id", "number")
plant_height_control_agg_1
```


### Cast aggregated table to get a column for each combination of culture and timepoint
```{r cast aggregated table}
plant_height_control_agg_cast <- cast(plant_height_control_agg, subspecies_id ~ culture_id + time_point)
head(plant_height_control_agg_cast)
dim(plant_height_control_agg_cast)
```


## Calculate normalized plant height values using ratio to median of culture at specific timepoint 
```{r normalized plant height}
# create matrix
plant_height_control_agg_cast_norm <- matrix(1:(6*34), ncol=6)
# run loop over all 6 columns in casted aggregated table
for (i in 1:6) {
  plant_height_control_agg_cast_norm[,i] <- plant_height_control_agg_cast[,i+1] / plant_height_control_agg_1$number[i]
}

# change colnames and rownames
colnames(plant_height_control_agg_cast_norm) <- colnames(plant_height_control_agg_cast)[2:7]
rownames(plant_height_control_agg_cast_norm) <- as.factor(plant_height_control_agg_cast$subspecies_id)
head(plant_height_control_agg_cast_norm)
dim(plant_height_control_agg_cast_norm)
```


## Melt normalized plant height table to get only 3 columns (all numbers in one column)
```{r melt normalized plant height table}
plant_height_control_agg_cast_norm_melt <- melt(plant_height_control_agg_cast_norm)

colnames(plant_height_control_agg_cast_norm_melt) <- c("subspecies_id", "culture_id_time_point", "number")

head(plant_height_control_agg_cast_norm_melt)
dim(plant_height_control_agg_cast_norm_melt)
```


## Visualize normalized plant height in heatmap and boxplot
```{r visualize normalized plant height in heatmap and boxplot}
par(oma=c(4,4,2,1))
heatmap.2(plant_height_control_agg_cast_norm, 
          col = bluered(75), trace = "none")
par(oma=c(0,0,0,0)) 

####################

# calculate median per subspecies
bymedian <- with(plant_height_control_agg_cast_norm_melt, reorder(subspecies_id, number, median))

# boxplot of plant height ordered by median per subspecies
boxplot(plant_height_control_agg_cast_norm_melt$number ~ bymedian, las=2)


# calculate mean per subspecies
bymean <- with(plant_height_control_agg_cast_norm_melt, reorder(subspecies_id, number, mean))

# boxplot of plant height ordered by median per subspecies
boxplot(plant_height_control_agg_cast_norm_melt$number ~ bymean, las=2)
```


## ANOVA for influence of factors on plant height
```{r anova for influence of factors on plant height}
# anova for influence of culture+timepoint
summary(aov(plant_height_control_agg_cast_norm_melt$number ~ plant_height_control_agg_cast_norm_melt$culture_id_time_point))

# anova for influence of subspecies
summary(aov(plant_height_control_agg_cast_norm_melt$number ~ plant_height_control_agg_cast_norm_melt$subspecies_id))
```


## Aggregate normalized plant height by subspecies --> MEAN (and median)
```{r aggregate normalized plant height by subspecies}
plant_height_control_agg_cast_norm_melt_agg <- aggregate(plant_height_control_agg_cast_norm_melt$number,
                                                         by=list(plant_height_control_agg_cast_norm_melt$subspecies_id),
                                                         mean)
colnames(plant_height_control_agg_cast_norm_melt_agg) <- c("subspecies_id", "mean")
dim(plant_height_control_agg_cast_norm_melt_agg)

plant_height_control_agg_cast_norm_melt_agg_median <- aggregate(plant_height_control_agg_cast_norm_melt$number,
                                                      by=list(plant_height_control_agg_cast_norm_melt$subspecies_id),
                                                      median)
colnames(plant_height_control_agg_cast_norm_melt_agg_median) <- c("subspecies_id", "median")

# sort list of normalized averaged plant height
plant_height_control_agg_cast_norm_melt_agg[order(plant_height_control_agg_cast_norm_melt_agg$mean), ]

# plot mean vs. median
plot(plant_height_control_agg_cast_norm_melt_agg$mean, plant_height_control_agg_cast_norm_melt_agg_median$median, pch=19)
abline(0,1, lty=3, lwd=2, col="red")

hist(plant_height_control_agg_cast_norm_melt_agg$mean, breaks=10, freq=FALSE, ylim=c(0,7))
lines(density(plant_height_control_agg_cast_norm_melt_agg$mean), col="green")
```


# Tolerance Classification
## Load tolerance data (new: from December 2013)

``gdrym131210labeled_export_R.txt`` comes from Karin and was uploaded to TROST webserver: http://trost.mpimp-golm.mpg.de/database/files/Karin108/gdrym131210labeled.csv
or http://trost.mpimp-golm.mpg.de/database/files/Karin108/gdrym131212.csv

```{r load tolerance data}
tolerance <- read.table("input/gdrym131210labeled_export_R.txt", header = T, sep = "\t")

# change colname
colnames(tolerance)[2] <- "subspecies_id"
head(tolerance)

# new drym values
summary(tolerance$mdrym_gh) # all greenhouse trials
summary(tolerance$mdrym_fve) # field trials with EARLY stress 
summary(tolerance$mdrym_fvl) # field trials with LATE stress 
summary(tolerance$mdrym_fva) # field trials AVERAGE 
```


### Take only tolerance data for GREENHOUSE (gh) and EARLY FIELD STRESS (fve) experiments
```{r plot DRYM values}
# histograms of DRYM values
par(mfrow=c(2,2))
hist(tolerance$mdrym_gh, xlim=c(-0.15, 0.2), col="lightblue", breaks=20, freq=FALSE, main="all greenhouse trials")
lines(density(tolerance$mdrym_gh), col="blue")

hist(tolerance$mdrym_fve, xlim=c(-0.15, 0.2), col="lightgreen", breaks=20, main="early field stress trials", ylim=c(0,10))
lines(density(tolerance$mdrym_fve), col="green")

# density plots
plot(density(tolerance$mdrym_fve), col="green", xlim=c(-0.2, 0.25))
lines(density(tolerance$mdrym_gh), col="blue")

# plot greenhouse vs. early field stress DRYM
plot(tolerance$mdrym_fve, tolerance$mdrym_gh, pch=19)
cor.test(tolerance$mdrym_fve, tolerance$mdrym_gh)
# p-value = 0.1116 --> no correlation

par(mfrow=c(1,1))
```


## Join tolerance data with phenotyper data (plant height and maturity group)
```{r join tolerance data with phenotyper data}
head(subspecies_query_result_part)
head(plant_height_control_agg_cast_norm_melt_agg)
head(tolerance)

phenotyper_metadata <- join(plant_height_control_agg_cast_norm_melt_agg, 
                            subspecies_query_result_part, 
                            by="subspecies_id")

tolerance_phenotyper_metadata_all <- join(phenotyper_metadata, tolerance, by="subspecies_id")
tolerance_phenotyper_metadata <- tolerance_phenotyper_metadata_all[, c(1,2,4,7,13,15)]

head(tolerance_phenotyper_metadata)
colnames(tolerance_phenotyper_metadata) <- c("subspecies_id", "norm_height_gh", "cultivar", "maturity", "mdrym_gh", "mdrym_fve")

write.table(tolerance_phenotyper_metadata, "output/tolerance_phenotyper_metadata.txt", sep="\t")
```


## Plot DRYM versus maturity and plant height
```{r plot DRYM versus maturity and plant height}
# drym greenhouse vs. maturity group
boxplot(tolerance_phenotyper_metadata$mdrym_gh ~ tolerance_phenotyper_metadata$maturity, 
        main="DRMY greenhouse", xlab="maturity class")

# drym field (early stress) vs. maturity group
boxplot(tolerance_phenotyper_metadata$mdrym_fve ~ tolerance_phenotyper_metadata$maturity, 
        main="DRMY field (early stress)", xlab="maturity class")

# group 5 has lower tolerance than 3  
# group 3 has lower tolerance than 2


# normalized plant height (greenhouse) vs. maturity group
boxplot(tolerance_phenotyper_metadata$norm_height_gh ~ tolerance_phenotyper_metadata$maturity,
        ylab = "norm height", xlab = "maturity class")

# drym greenhouse vs. normalized plant height (greenhouse)
plot(tolerance_phenotyper_metadata$mdrym_gh, tolerance_phenotyper_metadata$norm_height_gh)

# drym field (early stress) vs. normalized plant height (greenhouse)
plot(tolerance_phenotyper_metadata$mdrym_fve, tolerance_phenotyper_metadata$norm_height_gh)
```


## Select cultivars for training and test set of prediction model
### See Excel file for further information: ``D:\work\repos\trost_phenotypes\cultivar_selection_for_model.xlsx``
* Three cultivars are not used for modelling: Kolibri, Kiebitz and Kormoran (24, 25, 26)
* Set1 with 68% in training set and 32% in test set
* Set2 with 77% in training set and 23% in test set

```{r select cultivars}
train <- as.integer(c(382, 2673, 2674, 2675, 2853, 2854, 2855, 2857, 2858, 2860, 2863, 2864, 
                      2865, 2866, 2868, 2869, 2870, 2871, 2875, 2876, 2882)) #21 --> 68%
test <- as.integer(c(2856, 2859, 2861, 2862, 2867, 2877, 2878, 2879, 2880, 2881)) #10 --> 32%

train2 <- as.integer(c(382, 2673, 2674, 2675, 2853, 2854, 2855, 2857, 2858, 2860, 2863, 2864, 
                       2865, 2866, 2868, 2869, 2870, 2871, 2875, 2876, 2882, 2861, 2878, 2867)) #24 --> 77%
test2 <- as.integer(c(2856, 2859, 2862, 2877, 2879, 2880, 2881)) #7 --> 23%

train_data <- tolerance_phenotyper_metadata[which(tolerance_phenotyper_metadata$subspecies_id %in% train),]
test_data <- tolerance_phenotyper_metadata[which(tolerance_phenotyper_metadata$subspecies_id %in% test),]

train_data2 <- tolerance_phenotyper_metadata[which(tolerance_phenotyper_metadata$subspecies_id %in% train2),]
test_data2 <- tolerance_phenotyper_metadata[which(tolerance_phenotyper_metadata$subspecies_id %in% test2),]

# generate vector with test or training as value
model_set <- rep("NOT", 34)
model_set[which(tolerance_phenotyper_metadata$subspecies_id %in% train)] <- "train"
model_set[which(tolerance_phenotyper_metadata$subspecies_id %in% test)] <- "test"
model_set

model_set2 <- rep("NOT", 34)
model_set2[which(tolerance_phenotyper_metadata$subspecies_id %in% train2)] <- "train"
model_set2[which(tolerance_phenotyper_metadata$subspecies_id %in% test2)] <- "test"
model_set2

# add model set vector to table
tolerance_phenotyper_metadata_2 <- data.frame(tolerance_phenotyper_metadata,
                                              model_set)
head(tolerance_phenotyper_metadata_2)

tolerance_phenotyper_metadata_2_2 <- data.frame(tolerance_phenotyper_metadata, 
                                                model_set = model_set2)
head(tolerance_phenotyper_metadata_2_2)

# select only 31 cultivars, without Kolibri, Kormoran and Kiebitz
tolerance_phenotyper_metadata_2sub <- droplevels(subset(tolerance_phenotyper_metadata_2,
                                                        tolerance_phenotyper_metadata_2$model_set != "NOT"))

tolerance_phenotyper_metadata_2sub2 <- droplevels(subset(tolerance_phenotyper_metadata_2_2,
                                                         tolerance_phenotyper_metadata_2_2$model_set != "NOT"))
```


## Assess if training and test sets are balanced regarding:
* maturity
* normalized plant height (greenhouse, control)
* DRYM (gh, fve)

```{r assess balanced values}
# maturity
maturity_table <- table(tolerance_phenotyper_metadata_2sub$maturity,
                        tolerance_phenotyper_metadata_2sub$model_set)
maturity_table
prop.table(maturity_table, 1)
prop.table(maturity_table, 2)
assocplot(maturity_table)

# normalized plant height --> subsets for training and test set
norm_height_gh_train <- subset(tolerance_phenotyper_metadata_2sub$norm_height_gh, 
                               tolerance_phenotyper_metadata_2sub$model_set=="train")

norm_height_gh_test <- subset(tolerance_phenotyper_metadata_2sub$norm_height_gh, 
                              tolerance_phenotyper_metadata_2sub$model_set=="test")

# density plot of normalized plant height
plot(density(norm_height_gh_train), col="red", xlim=c(0.8,1.25), ylim=c(0,8))

lines(density(norm_height_gh_test))

hist(norm_height_gh_test, col="lightgreen", breaks=10)
```


## Plots with ggplots2 for DRYM and normalized plant height
```{r ggplots}
pdf("figures/compare_train_test_set.pdf")
qplot(model_set, norm_height_gh, data=tolerance_phenotyper_metadata_2sub, geom=c("boxplot", "jitter"),
      fill=model_set, main="normalized plant height \n (mean for all greenhouse trials)")

qplot(model_set, mdrym_gh, data=tolerance_phenotyper_metadata_2sub, geom=c("boxplot", "jitter"),
      fill=model_set, main="DRYM greenhouse trials")

qplot(model_set, mdrym_fve, data=tolerance_phenotyper_metadata_2sub, geom=c("boxplot", "jitter"),
      fill=model_set, main="DRYM field trials (early stress)")

ggplot(tolerance_phenotyper_metadata_2sub, aes(x=norm_height_gh, fill=model_set)) + 
  geom_density(alpha=.3) + scale_x_continuous(limits=c(0.8,1.25)) + ggtitle("normalized plant height \n (mean for all greenhouse trials)")

ggplot(tolerance_phenotyper_metadata_2sub, aes(x=mdrym_gh, fill=model_set)) + 
  geom_density(alpha=.3) + scale_x_continuous(limits=c(-0.2,0.25)) + ggtitle("DRYM greenhouse trials")

ggplot(tolerance_phenotyper_metadata_2sub, aes(x=mdrym_fve, fill=model_set)) + 
  geom_density(alpha=.3) + scale_x_continuous(limits=c(-0.15,0.2)) + ggtitle("DRYM field trials (early stress)")
dev.off()
```


## Convert drym to factor
### Convert drym to factor: Greenhouse
```{r convert drym to factor greenhouse}
# pot
summary(tolerance_phenotyper_metadata_2sub$mdrym_gh)
quantile(tolerance_phenotyper_metadata_2sub$mdrym_gh, probs=1/3) # upper third
quantile(tolerance_phenotyper_metadata_2sub$mdrym_gh, probs=2/3) # lower third
tolerance_phenotyper_metadata_2sub[which(tolerance_phenotyper_metadata_2sub$mdrym_gh < (-0.04245554)),] # low
tolerance_phenotyper_metadata_2sub[which(tolerance_phenotyper_metadata_2sub$mdrym_gh > 0.037723),] # high

tol_cat2_gh <- rep("low", 31)
tol_cat2_gh[which(tolerance_phenotyper_metadata_2sub$mdrym_gh>0)] <- "high"
table(tol_cat2_gh)

tol_cat3_gh <- rep("mid", 31)
tol_cat3_gh[which(tolerance_phenotyper_metadata_2sub$mdrym_gh<(-0.04245554))] <- "low"
tol_cat3_gh[which(tolerance_phenotyper_metadata_2sub$mdrym_gh>0.037723)] <- "high"
table(tol_cat3_gh)
```


### Convert drym to factor: Field
```{r convert drym to factor field}
# field
summary(tolerance_phenotyper_metadata_2sub$mdrym_fve)
quantile(tolerance_phenotyper_metadata_2sub$mdrym_fve, probs=1/3) # upper third
quantile(tolerance_phenotyper_metadata_2sub$mdrym_fve, probs=2/3) # lower third
tolerance_phenotyper_metadata_2sub[which(tolerance_phenotyper_metadata_2sub$mdrym_fve < (-0.01914815)),] # low
tolerance_phenotyper_metadata_2sub[which(tolerance_phenotyper_metadata_2sub$mdrym_fve > 0.02033081),] # high

tol_cat2_fve <- rep("low", 31)
tol_cat2_fve[which(tolerance_phenotyper_metadata_2sub$mdrym_fve>0)] <- "high"
table(tol_cat2_fve)

tol_cat3_fve <- rep("mid", 31)
tol_cat3_fve[which(tolerance_phenotyper_metadata_2sub$mdrym_fve<(-0.01914815))] <- "low"
tol_cat3_fve[which(tolerance_phenotyper_metadata_2sub$mdrym_fve>0.02033081)] <- "high"
table(tol_cat3_fve)
```


## Combine tolerance information in final table
```{r combine tolerance information in final table}
tolerance_phenotyper_metadata_2sub <- data.frame(tolerance_phenotyper_metadata_2sub, 
                                                 tol_cat2_gh, tol_cat3_gh, 
                                                 tol_cat2_fve, tol_cat3_fve)

tolerance_phenotyper_metadata_2sub2 <- data.frame(tolerance_phenotyper_metadata_2sub2, 
                                                  tol_cat2_gh, tol_cat3_gh, 
                                                  tol_cat2_fve, tol_cat3_fve)

table(tol_cat3_gh, tol_cat3_fve)
```


## Save final tolerance tables
* Set1 with 68% in training set and 32% in test set
* Set2 with 77% in training set and 23% in test set

```{r save final tolerance tables}
write.table(tolerance_phenotyper_metadata_2sub, 
            "output/tolerance_phenotyper_metadata_2sub.txt", sep="\t")

write.table(tolerance_phenotyper_metadata_2sub2, 
            "output/tolerance_phenotyper_metadata_2sub2.txt", sep="\t")
```


# Save workspace and sessionInfo
```{r save workspace}
save.image("tolerance_classification.RData")

sessionInfo()
```
