---
title: "Yield data analysis for Manuela"
author: "Heike Sprenger"
date: "Wednesday, July 13, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_phenotypes")
#setwd("~/work/repos/trost_phenotypes/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(ggplot2)
library(reshape)
library(pander)
library(multcomp)
library(extrafont)
library(plyr)
library(GGally)
library(gplots)
library(RColorBrewer)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
# load("yield_data_manuela.RData")
```


# Source R functions
```{r source R functions, include=FALSE}
source("../functions/names.R")
source("../functions/colors.R")
source("../functions/func_cv.R")
source("../functions/func_appendList.R")
source("../functions/func_anova_v2.R")
source("../functions/func_ttest.R")
source("../functions/func_aggregate_values.R")
source("../functions/func_modify_yield_data.R")
source("../functions/func_get_yield_data.R")
source("../functions/func_starch_yield_valdis.R")
source("../functions/func_calc_stress_index.R")
source("../functions/func_calc_relSY.R")
source("../functions/func_calc_relSY_table.R")
source("../functions/func_calc_drym.R")
source("../functions/func_calc_drym_table.R")
source("../functions/func_calc_ssi.R")
source("../functions/func_calc_ssi_table.R")
source("../functions/func_venn_diagram.R")
source("../functions/func_pairs_plot.R")

source("../functions/func_starch_yield_mpipruef.R")
source("../functions/func_starch_yield_feld.R")

#source("../functions/func_starch_yield_jki_feld_2012.R")
```


# Get yield data (from phenotyper)
```{r get yield data}
yield_data_trost <- func_get_yield_data(project = "trost")
yield_data_valdis <- func_get_yield_data(project = "valdis")

dim(yield_data_trost) # 34821    13
dim(yield_data_valdis) # 20347    14

table(yield_data_trost$culture)
table(yield_data_trost$attribute)
#table(yield_data_trost$culture, yield_data_trost$attribute)

# change colnames
colnames(yield_data_valdis)[3] <- "cultivar_id"
colnames(yield_data_valdis)[4] <- "cultivar"

write.table(yield_data_trost, "data/yield_data_manuela_raw.txt", sep="\t")
write.table(yield_data_valdis, "data/yield_data_manuela_valdis_raw.txt", sep="\t")
```


# Modify yield data
## Change alias for cultivar names
```{r change alias for cultivar names}
yield_data_trost$cultivar <- as.factor(yield_data_trost$cultivar)
levels(yield_data_trost$cultivar) <- names_cultivars_34

# VALDIS
idx_desiree <- which(grepl("^\\[St.D.n\\]", yield_data_valdis$name) | grepl("^\\[St.Desiree.n\\]", yield_data_valdis$name))
idx_ramses <- which(grepl("^\\[St.Ramses.n\\]", yield_data_valdis$name))
idx_euroresa <- which(grepl("^\\[St.Euroresa.n\\]", yield_data_valdis$name))
idx_albatros <- which(grepl("^\\[St.Albatros.n\\]", yield_data_valdis$name))

yield_data_valdis$cultivar[idx_desiree] <- "Desiree"
yield_data_valdis$cultivar[idx_ramses] <- "Ramses"
yield_data_valdis$cultivar[idx_euroresa] <- "Euroresa"
yield_data_valdis$cultivar[idx_albatros] <- "Albatros"

yield_data_valdis$cultivar <- as.factor(yield_data_valdis$cultivar)
table(yield_data_valdis$cultivar)

# remove plant_name column
yield_data_valdis <- yield_data_valdis[,-5]
```


## Merge Trost and Valdis data
```{r merge trost and valdis data}
yield_data <- rbind(yield_data_trost, yield_data_valdis)
dim(yield_data)
# 55168 13
```


## Change treatment name
```{r change treatment name}
yield_data$treatment_name <- yield_data$treatment
yield_data$treatment_name[which(yield_data$treatment == "169")] = "control"
yield_data$treatment_name[which(yield_data$treatment == "170")] = "drought stress"
yield_data$treatment_name[which(yield_data$treatment == "171")] = "control" # 50% nFK
yield_data$treatment[which(yield_data$treatment == "171")] = "169" # 50% nFK --> control

# remove entries with 30% nFK (id: 172, from Dethlingen)
idx_172 <- which(yield_data$treatment == "172")
yield_data <- yield_data[-idx_172,]
dim(yield_data)
# 53266 14

# change class of treatment_name
yield_data$treatment_name <- as.factor(yield_data$treatment_name)
levels(yield_data$treatment_name)

# table(yield_data$treatment_name)
# table(yield_data$treatment)
```


## Replace "-666.66" by NA
```{r replace -666.66 by NA}
# change class of number
yield_data$number <- as.numeric(yield_data$number)

idx_missing <- which(yield_data$number=="-666.66")
yield_data$number[idx_missing] <- NA
```


## Remove duplicate entries
```{r remove duplicate entries}
dim(yield_data)
# 53266 14

colnames(yield_data)

idx_duplicates <- which(duplicated(yield_data[,-c(6,7)])) # column 6, 7: measurement date and time should be ignored

yield_data_no_duplicates <- yield_data[-idx_duplicates,]
dim(yield_data_no_duplicates)
# 47260 14

write.table(yield_data_no_duplicates, "data/yield_data_manuela_no_duplicates.txt", sep="\t")
```


## Find outlier
```{r find outlier}
# outlier?
# idx_outlier <- which(yield_data_no_duplicates$number > 400)
# pander(yield_data_no_duplicates[idx_outlier,])
# yield_data_no_duplicates[idx_outlier, "plant_id"]

# remove them?
# entries with "999" as number belong to plants with FW = 0 --> change to NA
# yield_data_no_duplicates$number[idx_outlier] <- NA
```


# Get subset for specific cultures and cultivars
```{r get subset for specific cultures and cultivars}
table(yield_data$culture)
#table(yield_data$cultivar)

# without Dethlingen
# cultures_subset_trost <- c(56875, 62327, 45990, 57803, 44443, 56726, 62326, 56575, 58243, 60319, 62030) 

# with Dethlingen
cultures_subset_trost <- c(56875, 62327, 45990, 57803, 44443, 56726, 62326, 56575, 58243, 60319, 62030, 46150, 56877, 62328)

cultures_subset_valdis <- c(67199, 67516, 67518, 68015, 72237, 72247, 72275, 72292, 72396, 72482)

cultures_subset <- c(cultures_subset_trost, cultures_subset_valdis)

cultivars_subset <- c("Albatros", "Desiree", "Euroresa", "Ramses")
entities_subset <- c(12, 19, 810)

yield_subset <- droplevels(subset(yield_data_no_duplicates, yield_data_no_duplicates$culture %in% cultures_subset 
                                  & yield_data_no_duplicates$cultivar %in% cultivars_subset
                                  & yield_data_no_duplicates$entity_id %in% entities_subset))

table(yield_subset$culture, yield_subset$cultivar)

# contains unnecessary information
unwanted_entries <- which(yield_subset$entity_id == 810 & yield_subset$culture == 62326)
yield_subset <- yield_subset[-unwanted_entries,]
dim(yield_subset)
# 2912 14

length(unique(yield_subset$plant_id))
# 736

# change class of cultivar
yield_subset$cultivar <- as.factor(yield_subset$cultivar)
yield_subset$treatment <- as.factor(yield_subset$treatment)

levels(yield_subset$treatment)
levels(yield_subset$cultivar)
```


## Subset for plant height
* entity: "Spross" = 19
* value: "absolute Laenge" = 53
```{r subset for plant height}
plant_height <- droplevels(subset(yield_subset, yield_subset$entity_id == 19))

# remove plant height
yield_subset <- droplevels(subset(yield_subset, yield_subset$entity_id != 19))


table(plant_height$culture)
table(plant_height$measurement_date, plant_height$culture)

table(plant_height$measurement_date)
early <- c("2011-05-25", "2011-05-27", "2012-03-27", "2012-07-23", "2012-11-12")
late <- c("2011-06-16", "2012-04-24", "2012-06-18", "2012-08-22", "2012-12-10", "2012-12-11")

valdis_dates <- c("2014-04-30", "2014-05-07", "2014-05-12", "2014-05-21", "2014-05-26", "2014-05-27", 
                  "2014-05-30", "2014-06-03", "2014-06-10", "2014-06-11", "2015-04-27", "2015-05-04", 
                  "2015-05-06", "2015-05-12", "2015-05-13", "2015-05-18", "2015-05-20", "2015-05-27",
                  "2015-05-28", "2015-05-29", "2015-06-05", "2015-06-09", "2015-06-10", "2015-06-26")

plant_height_early <- subset(plant_height, plant_height$measurement_date %in% early)
plant_height_late <- subset(plant_height, plant_height$measurement_date %in% late)

# CONTAINS NO DATA! 
plant_height_valdis <- subset(plant_height, plant_height$measurement_date %in% valdis_dates)
```


## Calculate mean for replicated measurements
* MPI FGH 2014 and 2015 contain replicate measurements for starch contents -> calcualte mean!
```{r calculate mean for replicated measurements}
# remove wrong entries for JKI field trial 2012 (56875) from 2013-09-19
wrong_entries <- which(yield_subset$measurement_date == "2013-09-19" & yield_subset$culture == 56875)
yield_subset <- yield_subset[-wrong_entries,]

# summarize by calculating mean
yield_subset_correct <- ddply(yield_subset, c("plant_id", "culture", "cultivar_id", "cultivar", 
                                              "treatment", "measurement_date", 
                                              "entity_id", "entity_name",
                                              "attribute", "value_id", "unit", "treatment_name"), 
                          summarise, number = mean(as.numeric(number), na.rm = T))

dim(yield_subset) # 2112 14
dim(yield_subset_correct) # 2091 13

table(yield_subset$culture)
table(yield_subset_correct$culture)
```


## Calculate starch yield per experiment --> contains drymatter (%), starch content (g/kg) and starch yield (kg/plant and g/plant)
```{r calculate starch yield}
# TROST pot trials
pruef1 <- func_starch_yield_mpipruef(yield_subset, 56575)
pruef2 <- func_starch_yield_mpipruef(yield_subset, 58243)
pruef3 <- func_starch_yield_mpipruef(yield_subset, 60319)
pruef4 <- func_starch_yield_mpipruef(yield_subset, 62030)

jki_shelter_2011 <- func_starch_yield_feld(yield_subset, 45990, 1)
jki_shelter_2012 <- func_starch_yield_feld(yield_subset, 57803, 2)

# TROST field trials
mpi_field_2011 <- func_starch_yield_feld(yield_subset, 44443, 8)
mpi_field_2012 <- func_starch_yield_feld(yield_subset, 56726, 8)
mpi_field_2013 <- func_starch_yield_feld(yield_subset, 62326, 8)

jki_field_2012 <- func_starch_yield_feld(yield_subset, 56875, 6)
jki_field_2013 <- func_starch_yield_feld(yield_subset, 62327, 6)

dethlingen_2011 <- func_starch_yield_feld(yield_subset, 46150, 31)
dethlingen_2012 <- func_starch_yield_feld(yield_subset, 56877, 31)
dethlingen_2013 <- func_starch_yield_feld(yield_subset, 62328, 31)

# VALDIS
mpi_fgh_2014 <- func_starch_yield_feld(yield_subset_correct, 67199, 1)
mpi_fgh_2015 <- func_starch_yield_feld(yield_subset_correct, 72247, 1)

jki_shelter_2014 <- func_starch_yield_feld(yield_subset, 72237, 1)
jki_shelter_2015 <- func_starch_yield_feld(yield_subset, 72292, 1)

mpi_field_2014 <- func_starch_yield_feld(yield_subset, 67516, 5)
mpi_field_2015 <- func_starch_yield_feld(yield_subset, 72275, 5)

jki_field_2014 <- func_starch_yield_feld(yield_subset, 67518, 1)
jki_field_2015 <- func_starch_yield_feld(yield_subset, 72396, 5)

dethlingen_2015 <- func_starch_yield_feld(yield_subset, 72482, 10)
```


## Combine results
```{r combine results, fig.width=10}
starch_yield <- rbind(pruef1, pruef2, pruef3, pruef4,
                      jki_shelter_2011, jki_shelter_2012,
                      mpi_field_2012, mpi_field_2011, mpi_field_2013,
                      jki_field_2012, jki_field_2013,
                      dethlingen_2011, dethlingen_2012, dethlingen_2013,
                      mpi_fgh_2014, mpi_fgh_2015,
                      jki_shelter_2014, jki_shelter_2015,
                      mpi_field_2014, mpi_field_2015,
                      jki_field_2014, jki_field_2015,
                      dethlingen_2015)

dim(starch_yield)
# 736 10

# export table
write.table(starch_yield, "output/starch_yield_data_for_manu.txt", sep = "\t", quote = F, row.names = F)

length(unique(starch_yield$plant_id))
# 736

boxplot(starch_yield$starch_yield_g_per_plant ~ starch_yield$treatment * starch_yield$culture, 
        las = 2, col = c("blue", "red"))

# join starch yield with plant height
# starch_yield_joined <- join(starch_yield, plant_height_late, by="plant_id")
```



# Save workspace
```{r save workspace}
save.image("yield_data_manuela.RData")
```
